### 2017-06-24 Paper/paper.tex
Add TeX template

\documentclass[12pt,halfline,a4paper]{ouparticle}

\begin{document}

\title{Using a Two-Layer Neural Network and Physicochemical Properties to Classify Glass for Forensic Analysis}

\author{%
\name{Michael Arango, Mark Barna \& Paul Brewster}
}

\abstract{We need to figure out if we want an abstract in addition to the project proposal.....}

\date{\today}

\maketitle

\section{Project Proposal}
\label{sec1}

\section{Literature Review}
\label{sec2}

\section{Preparing your manuscript}
\label{sec3}

\subsection{General guidelines}
\label{sec3.1}

\begin{enumerate}
\item
\LaTeX\ and \AmS-\LaTeX\ provide a rich set of commands for all common, important
features of your paper. Use them; avoid definitions and use of custom commands.

\item
There is no need to redefine any \TeX, \LaTeX\ or \AmS-\LaTeX\ commands.

\item
Avoid direct formatting for headings cleanly set as section headings.

\item
Use \LaTeX\ commands for font changes. For example: use \verb+\textbf{phrase}+,
not \verb+{\bf phrase}+; use \verb+\mathcal{C}+, not \verb+{\cal C}+; etc.
\end{enumerate}

\subsection{How to start with \texttt{ouparticle.cls}}
\label{sec3.2}

Before you type anything that actually appears in the paper, you need to
include a \verb+\documentclass{ouparticle}+ command at the very beginning,
and then the two commands that have to be part of any \LaTeX\ document,
\verb+\begin{document}+ at the start and \verb+\end{document}+ at the
end of your paper.

\subsection{Document structure}
\label{sec3.3}

The main structure of your paper is as follows:

\begin{verbatim}
\documentclass[12pt,...]{ouparticle}
\usepackage[...]{packages}

\title{...}
\author{
    \name{...}
    \address{...}
    \email{...}
        \and
    \name{...}
    \address{...}
    \email{...}
        \and
    \name{...}
    \address{...}
    \email{...}
}
\abstract{...}
\keywords{...}

\maketitle

\begin{document}

\section{....}
...
\subsection{....}
....
\end{document}
\end{verbatim}

\subsection{Options}
\label{sec3.4}

By default, all of the options within \verb+article.cls+ are available
with this class file. This class file provides the following additional options.

\begin{description}
\item \textbf{oneline:}
This option will set your entire manuscript in one line spacing.
It will not affect the footnote, figure and table environments.

\item \textbf{halfline:}
This is to set your entire manuscript in half line spacing.

\item \textbf{endnotes:}
To make all footnotes to endnotes. You may follow the same
coding \verb+\footnote{text}+ for both footnotes and endnotes. Once you use this option
you have to use the \verb+\theendnotes+ command at the place where all the endnotes
have to be set in your paper.

\item \textbf{numbib:}
This is the default option that numbers the bibliography items;
this option does nothing with natbib and other packages.

\item \textbf{nonumbib:} For unnumbered bibliography.
\end{description}

\subsection{Front matter}
\label{sec3.5}

The title of the manuscript is simply specified by using the \verb+\title{text}+ command in
the same manner as in this sample. Author's information consists of the name of the author
and the corresponding institutions with addresses, as given in this example. Include an
electronic mail address if available, inserting it into the \verb+\email{text}+ commands.
You may follow the same coding if there are more than one author; separate authors with
\verb+\and+. Please identify the corresponding author with his/her electronic
mail address by \verb+\thanks{text}+. An abstract for your paper is specified by using
\verb+\abstract{text}+. A \verb+\keywords{text}+ macro may also be used to indicate keywords for the
article. Use \verb+\maketitle+ after the abstract and keywords to make the header of your article.

\subsection{Sections and subsections}
\label{sec3.6}

To begin a new section, give the heading of that section in the \verb+\section{text}+ command.
A section number is supplied automatically. Use the starred form (\verb+\section*{text}+) of the
command to suppress the automatic numbering. If you want to be able to make reference to that section,
then you need to \texttt{label} it (see Section \ref{sec3.14}). You can have sections up to
five levels. The sectioning commands are \verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
\verb|\paragraph| and \verb|\subparagraph|.

\subsection{Ordinary text}
\label{sec3.7}

The ends of words and sentences are marked by spaces. It does not matter how many
spaces you type. The end of a line counts as a space. One
or more blank lines denote the end of a paragraph.

There are a number of things for which you need to follow different
methods. As you know, quotation marks, quotes within quotes,
dashes, ellipsis, etc. should be as per the \LaTeX\ standard input. \LaTeX\ interprets some
common characters as commands, and therefore you must instead type those common characters as
specific \LaTeX\ commands to generate them. Those characters are \$, \&, \%, \#, \{, and \}.

\subsection{Formatting}
\label{sec3.8}

One should always use \LaTeX\ macros rather than the lower-level
\TeX\ macros like \verb+\it+, \verb+\bf+ and \verb+\tt+. The
\LaTeX\ macros offer much improved features. The following table summarizes the font
selection commands in \LaTeX.

\subsubsection*{\LaTeX\ text formatting commands}
\begin{tabular}{ll@{\hskip60pt}ll}
\verb+\textit+  & Italics      &\verb+\textsf+  & Sans Serif\\
\verb+\textbf+  & Boldface     &\verb+\textsc+  & Small Caps\\
\verb+\texttt+  & Typewriter   &\verb+\textmd+  & Medium Series\\
\verb+\textrm+  & Roman        &\verb+\textnormal+ & Normal Series\\
\verb+\textsl+  & Slanted      &\verb+\textup+  & Upright Series
\end{tabular}

\subsubsection*{\LaTeX\ math formatting commands}
\begin{tabular}{ll@{\qquad}ll}
\verb+\mathit+     & Math Italics            &\verb+\mathfrak+   & Fraktur\\
\verb+\mathbf+     & Math Boldface       &\verb+\mathbb+     & Blackboard Bold\\
\verb+\mathtt+     & Math Typewriter     &\verb+\mathnormal+ & Math Normal\\
\verb+\mathsf+     & Math Sans Serif     &\verb+\boldsymbol+ & Bold math for Greek letters\\
\verb+\mathcal+    & Calligraphic        &                   & and other symbols
\end{tabular}

\subsection{Figures and tables}
\label{sec3.9}

Use normal \LaTeX\ coding for figures and tables.
Figure and table environments should be inserted after (not in) the paragraph in which
the figure is first mentioned or grouped all
together at the end of the file. They will be numbered automatically.
The following is an example of typesetting a table.

\begin{verbatim}
\begin{table}
\caption{Table caption text.}
\label{key}
The table matter goes here.
\end{table}
\end{verbatim}

As always with \LaTeX, the \verb+\label+ must be after the
\verb+\caption+, and inside the figure or table environment. The reference for
figures and tables inside text can be made using the \verb|\ref{key}| command.

\subsection{Equations}\label{sec3.10}

Equations are used in the same way as described in the \LaTeX\ manual.
Do not start a paragraph with a displayed equation. Equations are numbered consecutively, with equation numbers
in parentheses flush right.
 For example, if you type
\begin{verbatim}
\begin{equation}\label{eq1}
\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = [\sigma r_2/(2\mu_0)]
\int^{\infty}_0\exp(-\lambda|z_j-z_i|)\lambda^{-1}J_1 (\lambda r_2)J_0
(\lambda r_i\,\lambda {\rm d}\lambda)
\end{equation}
\end{verbatim}
then you will get the following output:
\begin{equation}\label{eq1}
\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = [\sigma r_2/(2\mu_0)]\int^{\infty}_0
\exp(-\lambda|z_j-z_i|)\lambda^{-1}J_1 (\lambda r_2)J_0 (\lambda r_i\,\lambda {\rm d}\lambda)
\end{equation}
It inserts space both above and below the equation. \AmS-\LaTeX{} has several environments that
make it easier to typeset complicated multiline displayed equations. These are explained in the
\AmS-\LaTeX{} User Guide. A \verb+subequation+ environment is available to create equations with
sub-numbering of the equation counter. It takes one (optional)
argument to specify the way that the sub-counter should appear.

\subsection{Displayed text}
\label{sec3.11}

Text is displayed by indenting it from the left and right margins.
Quotations are commonly displayed. There are short
quotations:
\begin{quote}
   This is a short quotation.  It consists of a
   single paragraph of text.  See how it is formatted.
\end{quote}
and longer ones:
\begin{quotation}
   This is a longer quotation.  It consists of two
   paragraphs of text, neither of which are
   particularly interesting.

   This is the second paragraph of the quotation.  It
   is just as dull as the first paragraph.
\end{quotation}
You can even display poetry.
\begin{verse}
   There is an environment
    for verse \\             % The \\ command separates lines
   Whose features some poets % within a stanza.
   will curse.

                             % One or more blank lines separate stanzas.

   For instead of making\\
   Them do \emph{all} line breaking, \\
   It allows them to put too many words on a line when they'd rather be
   forced to be terse.
\end{verse}

\subsection{Listings}
\label{sec3.12}

Another frequently displayed structure is a list. The
following is an example of an \emph{itemized} list.
\begin{itemize}
   \item This is the first item of an itemized list.
         Each item in the list is marked with a
         `$\bullet$'.

   \item This is the second item of the list. It
         contains another list nested inside it. The
         inner list is an \emph{enumerated} list.
         \begin{enumerate}
            \item This is the first item of an enumerated
                  list that is nested within the
                  itemized list.

            \item This is the second item of the inner list.
                  \LaTeX\ allows you to nest lists
                  deeper than you really should.
         \end{enumerate}
         This is the rest of the second item of the
         outer list. It is no more interesting than
         any other part of the item.
   \item This is the third item of the list.
\end{itemize}

\subsection{Displayed sentences: theorems and such}
\label{sec3.13}

These environments have to be defined with the help of \LaTeX's \verb+\newtheorem+ command, and
also with the \AmS-\LaTeX\ package for theorems that is already with your class file.
For example, \verb+\newtheorem{thm}{Theorem}+. Predefined theorem styles can be used in your article
to differentiate the theorem-like environments. You can have an extra command, \verb+\newproof+,
that can be used for displayed text. The following is an example of using the above-defined
\verb+thm+ environment.
\begin{verbatim}
\begin{thm}
This is body matter for this environment.
\end{thm}
\end{verbatim}

\subsection{Cross-referencing}
\label{sec3.14}

\LaTeX\ possesses features for labelling and cross-referencing
section headings, equations, tables, figures and theorems.
Their proper usage in the context of section headings, equations,
tables and figures are discussed in the appropriate sections.

Cross-referencing depends upon the use of `keys' that are defined by the user.
The \verb+\label{key}+ command is used to identify the links. Keys are strings of
characters that serve to label section headings, equations, tables and figures
that replace explicit, by-hand numbering. The \verb+\ref{key}+ command is used for
cross-referencing.

Files that use cross-referencing (and almost all manuscripts do)
need to be processed through \LaTeX\ at least twice to
ensure that the keys have been properly linked to the appropriate numbers.

\subsection{Footnotes and endnotes}
\label{sec3.15}

The footnote text can either appear at the bottom of a page or at the end of your paper.
The \verb+\footnote+ macro \emph{should not} be used in the front matter to provide additional
information about authors (such as corresponding addresses); instead, use \verb+\thanks{text}+ commands.
The document option `\texttt{endnotes}' is used to make endnotes. The command \verb+\theendnotes+ should
be used to place the endnotes at the required location in the text. They will be put in a separate
`Notes' section.

\subsection{Appendix}
\label{sec3.16}

The \verb+\appendix+ command signals that all following sections are
appendices, and therefore the headings after \verb+\appendix+ will be set
as appendix headings. For a single appendix, use \verb+\appendix*+ followed by the \verb+\section{text}+
command to suppress the appendix letter in the section heading.

\subsection{Special sections for notes and acknowledgements}
\label{sec3.17}

If you wish to include a `Notes' or `Acknowledgements' section in your paper,
use the \verb+\begin{notes}...\end{notes}+ macro. We use the same environment for both
`Notes' and `Acknowledgements'. The following examples show to how to use this macro.
\begin{verbatim}
\begin{notes}
Please note that this class file is provided as it is, and
copyright by Oxford University Press. You are free to use this
class file, provided that you do not make changes in this class file.
If you do make changes, you are requested to rename the class file.
\end{notes}

\begin{notes}[Acknowledgements]
The authors would like to thank...
\end{notes}

\end{verbatim}

\subsection{References}
\label{sec3.18}

The reference entries can be \LaTeX\ typed bibliographies or generated through a BIB\TeX\ database.
BIB\TeX\ is an adjunct to \LaTeX\ that aids in the preparation of bibliographies. BIB\TeX\
allows authors to build up a database or collection of bibliography entries that may be used for many
manuscripts. They also save us the trouble of having to specify formatting. More details can be found
in the \textit{BIB\TeX\ Guide}. For \LaTeX\ reference entries use the
\verb+\begin{thebibliography}....\end{thebibliography}+ environment (see below) to make references in your paper.
We have provided the class file option to distinguish two styles of references. Those options are \verb+numbib+ and \verb+nonumbib+.
You can select one of these options with the \verb+\documentclass+ command. By default the class file will take the
\verb+numbib+ option. The following is an example of \LaTeX\ bibliography.

\begin{verbatim}
\begin{thebibliography}{0}
\bibitem{bib1}
Goossens, M., F. Mittelbach, and A. Samarin: {\em The {\LaTeX} Companion}.
Addison-Wesley, Reading, MA, USA, 1994.
\bibitem{bib2}
Knuth, D.E: {\em The {\TeX}book}. Addison-Wesley, Reading, MA, USA, 1984.
\bibitem{bib3}
Lamport, L.: {\em {\LaTeX} -- A Document Preparation System -- User's
Guide and Reference Manual}. Addison-Wesley, Reading, MA, USA, 1985.
\bibitem{bib4}
Smith, I.N., R.S. Johnes, and W.P. Hines: 1992, `Title of the Article',
\textit{Journal Title in Italics} \textbf{Vol. no. X}, pp. 00--00
\end{thebibliography}
\end{verbatim}

\section{Macro packages}
\label{sec4}

The following packages are compulsorily needed by the class file:
\begin{verbatim}
amsmath         graphicx
amssymb         endnotes
amsfonts        setspace
verbatim        geometry
\end{verbatim}

The commonly used packages already used by this class file that authors can use whenever required are:
\begin{verbatim}
xspace          latexsym        url
amscd           multicol        algorithm
rotating        array           subfigure
\end{verbatim}

Additionally, you can use other packages and these should be loaded
using the \verb+\usepackage+ command.

\end{document}

### 2017-06-24 neuron.py
split dataset
In addition to creating the split function, I also moved the jacobian
matrix and sensitivity calculations outside of main() to their own
functions. I haven’t yet coded how the program handles the validation
set but believe it should be part of the main loop iteration somehow.

@@ -1,6 +1,6 @@
 import numpy as np
+import pandas as pd
 import matplotlib.pyplot as plt
-
 
 # create neuron class
 class Neuron:
@@ -41,21 +41,48 @@ class Neuron:
         a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
         return a
 
+    # softmax derivative jacobian matrix
+    def j_softmax(self,a,n):
+        jacobian = np.empty([n, n])
+        for i in range(n):
+            for j in range(n):
+                if i == j:
+                    jacobian[i, j] = a[i] * np.sum(a - a[i])
+                else:
+                    jacobian[i, j] = -a[i] * a[j]
+        return jacobian
+
     # hyperbolic tangent sigmoid transfer function
     def tangsig(self):
         a = 2/(1 + np.exp(-2 * self.net_input())) - 1
         return a
 
-
-# define target function to approximate
-def target(p):
-    g = 1 + np.sin(np.pi/2*p)
-    return g
+    # tansig derivative jacobian matrix
+    def j_tansig(self,a,n):
+        jacobian = np.diag(np.ones(n)) - np.diagflat(np.power(a, 2))
+        return jacobian
 
 # define error calculation function
 def error_calc(output, target):
     e = target - output
     return e
+
+# define cross entropy error calculation function for softmax (assuming targets of 1,0)
+def cross_entropy(a):
+    i = np.where(a == 0)
+    a[i] = 1e-15 # replace 0s so ln doesn't produce infinity
+    e = -np.log(a).sum()
+    return e
+
+# calculate hidden layer sensitivity
+def senseh(F_prime,W_1,s_1):
+    s = F_prime * np.transpose(W_1) * s_1
+    return s
+
+# calculate output layer sensitivity
+def senseo(F_prime,e):
+    s = -2 * F_prime * e
+    return s
 
 # define learning rule function
 def learn(weight_old, bias_old, sensitivity, input, learning_rate):
@@ -63,15 +90,50 @@ def learn(weight_old, bias_old, sensitivity, input, learning_rate):
     bias_new = bias_old - learning_rate * sensitivity
     return  weight_new, bias_new
 
+# split dataset into training, validation, testing sets
+def split(dataset,t,v):
+    # split dataset by target class
+    dataset0 = dataset.ix[dataset.ix[:,'target']==0,:]
[…]

### 2017-06-24 neuron.py
Update to split function
Updated split function to remove the old index column.

@@ -115,7 +115,7 @@ def split(dataset,t,v):
     test = test.sample(frac=1)
     test.reset_index(inplace=True)
 
-    return train, validate, test
+    return train.iloc[:,1:], validate.iloc[:,1:], test.iloc[:,1:]
 
 def main():
 
@@ -149,7 +149,7 @@ def main():
     for epoch in range(iterations):
 
         # first layer
-        neuron1 = Neuron(input=train[epoch],weight=W1,bias=b1)
+        neuron1 = Neuron(input=train.iloc[epoch,:-1],weight=W1,bias=b1)
         a1 = neuron1.tangsig()
         # second layer
         neuron2 = Neuron(input=a1,weight=W2,bias=b2)
@@ -169,12 +169,14 @@ def main():
         # calculate new weight and bias for layer 1
         W1, b1 = learn(weight_old=W1,bias_old=b1,sensitivity=s1,input=p_sample,learning_rate=alpha)
 
+avg_cost_t = np.mean(cost_t)
+
     # test network
     neuron1 = Neuron(input=test,weight=W1,bias=b1)
     a1 = neuron1.tangsig()
     neuron2 = Neuron(input=a1,weight=W2,bias=b2)
     a2 = neuron2.softmax()
-    
+
 
 if __name__ == '__main__':
     main()

### 2017-06-24 neuron.py
Fixed W1 initialization
Initialize W1 to include number of input features.

@@ -130,13 +130,13 @@ def main():
     train, validate, test = split(glass, t=0.7, v=0.15)
 
     # specify number of neurons in each layer and the learning rate
-    num_neurons1 = 10
+    num_neurons1 = 2
     num_neurons2 = 2
     alpha = 0.2
     iterations = 100
 
     # initialize weight and bias randomly for each layer from -0.5 to 0.5
-    W1 = np.matrix(np.random.rand(num_neurons1,1) - 0.5)
+    W1 = np.matrix(np.random.rand(num_neurons1,train.iloc[0,:-1].count()) - 0.5)
     b1 = np.matrix(np.random.rand(num_neurons1,1) - 0.5)
     W2 = np.matrix(np.random.rand(num_neurons2,num_neurons1) - 0.5)
     b2 = np.matrix(np.random.rand(num_neurons2,1) - 0.5)
@@ -168,8 +168,9 @@ def main():
         W2, b2 = learn(weight_old=W2,bias_old=b2,sensitivity=s2,input=a1,learning_rate=alpha)
         # calculate new weight and bias for layer 1
         W1, b1 = learn(weight_old=W1,bias_old=b1,sensitivity=s1,input=p_sample,learning_rate=alpha)
+        pass
 
-avg_cost_t = np.mean(cost_t)
+    avg_cost_t = np.mean(cost_t)
 
     # test network
     neuron1 = Neuron(input=test,weight=W1,bias=b1)

### 2017-06-24 neuron.py
Fixed input matrix shape

@@ -148,8 +148,13 @@ def main():
     # training & validation
     for epoch in range(iterations):
 
+        # create input matrix from training dataset with number of columns to match neurons
+        input = np.matrix(train.iloc[epoch,:-1])
+        input = np.concatenate([input]*num_neurons1)
+        input = np.transpose(input)
+
         # first layer
-        neuron1 = Neuron(input=train.iloc[epoch,:-1],weight=W1,bias=b1)
+        neuron1 = Neuron(input=input,weight=W1,bias=b1)
         a1 = neuron1.tangsig()
         # second layer
         neuron2 = Neuron(input=a1,weight=W2,bias=b2)

### 2017-06-24 neuron.py
corrected input matrix
Corrected input matrix duplicate row

@@ -150,7 +150,7 @@ def main():
 
         # create input matrix from training dataset with number of columns to match neurons
         input = np.matrix(train.iloc[epoch,:-1])
-        input = np.concatenate([input]*num_neurons1)
+        #input = np.concatenate([input]*num_neurons1)
         input = np.transpose(input)
 
         # first layer

### 2017-06-24 Paper/paper.tex
Uploading Project Proposal

@@
 \author{%
-\name{Michael Arango, Mark Barna \& Paul Brewster}
-}
+\name{Michael Arango, Mark Barna \& Paul Brewster}}
 
-\abstract{We need to figure out if we want an abstract in addition to the project proposal.....}
-
+%\abstract{abstract text goes here...}
 \date{\today}
-
 \maketitle
@@
 \label{sec1}
+In this study, we will use the physicochemical properties of glass to determine whether or not a given glass sample was taken from a window.
+This is a fundamental problem in forensic analysis as it is highly unlikely that glass fragments will be found on people unless they have been present at the time glass breaks.
+Glass analysis is of vital importance in forensic science as it allows us to test if the glass fragment found on a person is the same as the glass at a crime scene.
+Since glass is made up of several raw materials and certain elements impart specific properties, we can find out a lot about the glass if we analyze the chemical composition.
+
+The dataset we chose for analysis was made available for download from the UCI Machine Learning Repository and was created by the USA Forensic Science Service.
+There are 214 observations of 9 different features along with 214 targets that specify whether the glass sample came from a window or not.
+While we would like more data to train a neural network, we believe the dataset is large enough for our purposes.
+It is difficult to know before we train a neural network if we have enough data, but the amount of data required is directly related to the complexity of the underlying decision boundary we are trying to implement.
+We won't
+know how complex the decision boundary we are trying to approximate is until we train the network, but we feel confident using the dataset as many others have used the dataset and found robust results.
+Several others papers in the literature use much more complex methods than we will employ and have not found the size of the data to be an issue.
+
+We have chosen a two-layer perceptron network with tangent-sigmoid transfer functions in the hidden layer and \emph{softmax} transfer functions in the output layer.
+This is a fairly standard network for pattern recognition.
+Moreover, we will use the \emph{Scaled Conjugate Gradient (SGD)} algorithm to train the network as it is good for pattern recognition problems in which the output layer uses a non-linear transfer function.
+Since we do not expect the training error to converge to zero, we implement early stopping criteria to prevent overfitting.
+Lastly, we use \emph{cross-entropy} as our performance index since our targets take on discrete values and it is the optimal performance index for pattern recognition networks that use the \emph{softmax} transfer function in the output layer.
+
+Two different frameworks will be used to implement the neural network.
+First, we will use the Neural Network Toolbox, specifically the Neural Network Pattern Recognition Tool (\verb|nprtool|) train, validate, and test our network.
+We use this framework to start with a simple graphical user interface to quickly ensure our specified network architecture is appropriate and to get baseline performance statistics.
+Then, we will replicate the analysis in Python to gain practical experience building network architectures in a scripting language.
+Note that since the goal is practical experience, we will not be leveraging the power of the \emph{scikit-learn} package (\verb|sklearn|) for this exercise.
+
+Several reference materials will be consulted to obtain sufficient background knowledge of the subject at hand.
+First, we plan on doing a thorough review of the forensic chemistry and geology literature to understand the reasons for using physicochemical properties to classify glass.
+Then, papers on glass analysis will be examined to supplement background knowledge with experiential knowledge.
+
+Considering our problem is one of pattern recognition, a confusion matrix will be used to assess the accuracy of our model and the \emph{false postive} (Type I error) and \emph{false negative} (Type II error) rates.
+Further, the \emph{Receiver Operating Characteristic (ROC) curve} will be used to compare the true positive rate to the false positive rate.
+This will help us gain additional knowledge of the predictive power of our network.
+
+We plan to finish our research and submit it by Wednesday, June 28, 2017.
 

### 2017-06-25 Paper/paper.tex
Paper Editing and Formatting

@@
 \author{%
-\name{Michael Arango, Mark Barna \& Paul Brewster}}
+\name{Michael Arango}
+\email{mikearango@gwu.edu}
+	\and
+\name{Mark Barna}
+\email{mark.barna@gmail.com}
+	\and
+\name{Paul Brewster}
+\email{pfbrewster@gmail.com}
+}
 
@@
 \maketitle
+\pagebreak
+\tableofcontents
+\pagebreak
 
@@
 know how complex the decision boundary we are trying to approximate is until we train the network, but we feel confident using the dataset as many others have used the dataset and found robust results.
-Several others papers in the literature use much more complex methods than we will employ and have not found the size of the data to be an issue.
+Several other papers in the literature use much more complex methods than we will employ and have not found the size of the data to be an issue.
 
@@
 
-\section{Literature Review}
+\section{Introduction}
 \label{sec2}
+An overview of the project and an outline of the report (I like to write intro after I finish a project).
 
-\section{Preparing your manuscript}
+\subsection{Literature Review}
 \label{sec3}
+
+\section{Description of the Dataset}
+\subsection{Inputs}
+\subsection{Targets}
+
+\section{Description of the Network Architecture and Training Algorithm}
+
+\subsection{Network Architecture}
+\subsection{Training Algorith}
+
+\section{Experimental Setup}
+
+\subsection{Data Preprocessing}
+\subsection{Implementation of the Network}
+\subsection{Performance Index}
+
+\section{Results}
+
+\section{Conclusion}
 
@@
 \label{sec3.1}
-
-\begin{enumerate}
-\item
-\LaTeX\ and \AmS-\LaTeX\ provide a rich set of commands for all common, important
-features of your paper. Use them; avoid definitions and use of custom commands.
-
-\item
-There is no need to redefine any \TeX, \LaTeX\ or \AmS-\LaTeX\ commands.
-
-\item
-Avoid direct formatting for headings cleanly set as section headings.
-
-\item
-Use \LaTeX\ commands for font changes. For example: use \verb+\textbf{phrase}+,
-not \verb+{\bf phrase}+; use \verb+\mathcal{C}+, not \verb+{\cal C}+; etc.
-\end{enumerate}
-
-\subsection{How to start with \texttt{ouparticle.cls}}
-\label{sec3.2}
-
-Before you type anything that actually appears in the paper, you need to
-include a \verb+\documentclass{ouparticle}+ command at the very beginning,
-and then the two commands that have to be part of any \LaTeX\ document,
-\verb+\begin{document}+ at the start and \verb+\end{document}+ at the
-end of your paper.
-
-\subsection{Document structure}
-\label{sec3.3}
-
-The main structure of your paper is as follows:
-
-\begin{verbatim}
-\documentclass[12pt,...]{ouparticle}
-\usepackage[...]{packages}
-
-\title{...}
-\author{
-    \name{...}
-    \address{...}
-    \email{...}
-        \and
-    \name{...}
-    \address{...}
-    \email{...}
-        \and
-    \name{...}
-    \address{...}
-    \email{...}
-}
-\abstract{...}
-\keywords{...}
-
-\maketitle
-
-\begin{document}
-
-\section{....}
-...
-\subsection{....}
-....
-\end{document}
-\end{verbatim}
-
-\subsection{Options}
-\label{sec3.4}
 
@@
 \begin{description}
-\item \textbf{oneline:}
-This option will set your entire manuscript in one line spacing.
-It will not affect the footnote, figure and table environments.
-
-\item \textbf{halfline:}
-This is to set your entire manuscript in half line spacing.
 
@@
 
-\item \textbf{nonumbib:} For unnumbered bibliography.
 \end{description}
-
-\subsection{Front matter}
-\label{sec3.5}
-
-The title of the manuscript is simply specified by using the \verb+\title{text}+ command in
-the same manner as in this sample. Author's information consists of the name of the author
-and the corresponding institutions with addresses, as given in this example. Include an
-electronic mail address if available, inserting it into the \verb+\email{text}+ commands.
-You may follow the same coding if there are more than one author; separate authors with
-\verb+\and+. Please identify the corresponding author with his/her electronic
-mail address by \verb+\thanks{text}+. An abstract for your paper is specified by using
-\verb+\abstract{text}+. A \verb+\keywords{text}+ macro may also be used to indicate keywords for the
-article. Use \verb+\maketitle+ after the abstract and keywords to make the header of your article.
 
@@
 
-To begin a new section, give the heading of that section in the \verb+\section{text}+ command.
-A section number is supplied automatically. Use the starred form (\verb+\section*{text}+) of the
-command to suppress the automatic numbering. If you want to be able to make reference to that section,
-then you need to \texttt{label} it (see Section \ref{sec3.14}). You can have sections up to
-five levels. The sectioning commands are \verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
-\verb|\paragraph| and \verb|\subparagraph|.
-
-\subsection{Ordinary text}
-\label{sec3.7}
-
-The ends of words and sentences are marked by spaces. It does not matter how many
-spaces you type. The end of a line counts as a space. One
-or more blank lines denote the end of a paragraph.
-
-There are a number of things for which you need to follow different
-methods. As you know, quotation marks, quotes within quotes,
-dashes, ellipsis, etc. should be as per the \LaTeX\ standard input. \LaTeX\ interprets some
-common characters as commands, and therefore you must instead type those common characters as
-specific \LaTeX\ commands to generate them. Those characters are \$, \&, \%, \#, \{, and \}.
-
-\subsection{Formatting}
-\label{sec3.8}
-
-One should always use \LaTeX\ macros rather than the lower-level
-\TeX\ macros like \verb+\it+, \verb+\bf+ and \verb+\tt+. The
-\LaTeX\ macros offer much improved features. The following table summarizes the font
-selection commands in \LaTeX.
+ If you want to be able to make reference to that section,
+then you need to \texttt{label} it (see Section \ref{sec3.14}).
 
@@
 
-Use normal \LaTeX\ coding for figures and tables.
-Figure and table environments should be inserted after (not in) the paragraph in which
-the figure is first mentioned or grouped all
-together at the end of the file. They will be numbered automatically.
 The following is an example of typesetting a table.
@@
 
-Equations are used in the same way as described in the \LaTeX\ manual.
-Do not start a paragraph with a displayed equation. Equations are numbered consecutively, with equation numbers
-in parentheses flush right.
 For example, if you type
@@
 \end{equation}
-It inserts space both above and below the equation. \AmS-\LaTeX{} has several environments that
-make it easier to typeset complicated multiline displayed equations. These are explained in the
-\AmS-\LaTeX{} User Guide. A \verb+subequation+ environment is available to create equations with
-sub-numbering of the equation counter. It takes one (optional)
-argument to specify the way that the sub-counter should appear.
-
-\subsection{Displayed text}
-\label{sec3.11}
-
-Text is displayed by indenting it from the left and right margins.
-Quotations are commonly displayed. There are short
-quotations:
-\begin{quote}
-   This is a short quotation.  It consists of a
-   single paragraph of text.  See how it is formatted.
-\end{quote}
-and longer ones:
-\begin{quotation}
-   This is a longer quotation.  It consists of two
-   paragraphs of text, neither of which are
-   particularly interesting.
-
-   This is the second paragraph of the quotation.  It
-   is just as dull as the first paragraph.
-\end{quotation}
-You can even display poetry.
-\begin{verse}
-   There is an environment
-    for verse \\             % The \\ command separates lines
-   Whose features some poets % within a stanza.
-   will curse.
-
-                             % One or more blank lines separate stanzas.
-
-   For instead of making\\
-   Them do \emph{all} line breaking, \\
-   It allows them to put too many words on a line when they'd rather be
-   forced to be terse.
-\end{verse}
 
@@
 
-\subsection{Displayed sentences: theorems and such}
-\label{sec3.13}
+\pagebreak
 
-These environments have to be defined with the help of \LaTeX's \verb+\newtheorem+ command, and
-also with the \AmS-\LaTeX\ package for theorems that is already with your class file.
-For example, \verb+\newtheorem{thm}{Theorem}+. Predefined theorem styles can be used in your article
-to differentiate the theorem-like environments. You can have an extra command, \verb+\newproof+,
-that can be used for displayed text. The following is an example of using the above-defined
-\verb+thm+ environment.
-\begin{verbatim}
-\begin{thm}
-This is body matter for this environment.
-\end{thm}
-\end{verbatim}
-
-\subsection{Cross-referencing}
-\label{sec3.14}
-
-\LaTeX\ possesses features for labelling and cross-referencing
-section headings, equations, tables, figures and theorems.
-Their proper usage in the context of section headings, equations,
-tables and figures are discussed in the appropriate sections.
-
-Cross-referencing depends upon the use of `keys' that are defined by the user.
-The \verb+\label{key}+ command is used to identify the links. Keys are strings of
-characters that serve to label section headings, equations, tables and figures
-that replace explicit, by-hand numbering. The \verb+\ref{key}+ command is used for
-cross-referencing.
-
-Files that use cross-referencing (and almost all manuscripts do)
-need to be processed through \LaTeX\ at least twice to
-ensure that the keys have been properly linked to the appropriate numbers.
-
-\subsection{Footnotes and endnotes}
-\label{sec3.15}
-
-The footnote text can either appear at the bottom of a page or at the end of your paper.
-The \verb+\footnote+ macro \emph{should not} be used in the front matter to provide additional
-information about authors (such as corresponding addresses); instead, use \verb+\thanks{text}+ commands.
-The document option `\texttt{endnotes}' is used to make endnotes. The command \verb+\theendnotes+ should
-be used to place the endnotes at the required location in the text. They will be put in a separate
-`Notes' section.
-
-\subsection{Appendix}
-\label{sec3.16}
-
-The \verb+\appendix+ command signals that all following sections are
-appendices, and therefore the headings after \verb+\appendix+ will be set
-as appendix headings. For a single appendix, use \verb+\appendix*+ followed by the \verb+\section{text}+
-command to suppress the appendix letter in the section heading.
-
-\subsection{Special sections for notes and acknowledgements}
-\label{sec3.17}
-
-If you wish to include a `Notes' or `Acknowledgements' section in your paper,
-use the \verb+\begin{notes}...\end{notes}+ macro. We use the same environment for both
-`Notes' and `Acknowledgements'. The following examples show to how to use this macro.
-\begin{verbatim}
-\begin{notes}
-Please note that this class file is provided as it is, and
-copyright by Oxford University Press. You are free to use this
-class file, provided that you do not make changes in this class file.
-If you do make changes, you are requested to rename the class file.
-\end{notes}
-
-\begin{notes}[Acknowledgements]
-The authors would like to thank...
-\end{notes}
-
-\end{verbatim}
-
-\subsection{References}
-\label{sec3.18}
-
-The reference entries can be \LaTeX\ typed bibliographies or generated through a BIB\TeX\ database.
-BIB\TeX\ is an adjunct to \LaTeX\ that aids in the preparation of bibliographies. BIB\TeX\
-allows authors to build up a database or collection of bibliography entries that may be used for many
-manuscripts. They also save us the trouble of having to specify formatting. More details can be found
-in the \textit{BIB\TeX\ Guide}. For \LaTeX\ reference entries use the
-\verb+\begin{thebibliography}....\end{thebibliography}+ environment (see below) to make references in your paper.
-We have provided the class file option to distinguish two styles of references. Those options are \verb+numbib+ and \verb+nonumbib+.
-You can select one of these options with the \verb+\documentclass+ command. By default the class file will take the
-\verb+numbib+ option. The following is an example of \LaTeX\ bibliography.
-
-\begin{verbatim}
+\addcontentsline{toc}{section}{References}
 \begin{thebibliography}{0}
@@
 \end{thebibliography}
-\end{verbatim}
 
-\section{Macro packages}
-\label{sec4}
+\pagebreak
+\appendix
 
-The following packages are compulsorily needed by the class file:
-\begin{verbatim}
-amsmath         graphicx
-amssymb         endnotes
-amsfonts        setspace
-verbatim        geometry
-\end{verbatim}
-
-The commonly used packages already used by this class file that authors can use whenever required are:
-\begin{verbatim}
-xspace          latexsym        url
-amscd           multicol        algorithm
-rotating        array           subfigure
-\end{verbatim}
-
-Additionally, you can use other packages and these should be loaded
-using the \verb+\usepackage+ command.
+\section{Perceptron}
+\section{Backprop}
+\section{Subroutine X}
 

### 2017-06-25 neuron.py
Updated cross entropy function

@@ -62,16 +62,11 @@ class Neuron:
         jacobian = np.diag(np.ones(n)) - np.diagflat(np.power(a, 2))
         return jacobian
 
-# define error calculation function
-def error_calc(output, target):
-    e = target - output
-    return e
-
 # define cross entropy error calculation function for softmax (assuming targets of 1,0)
-def cross_entropy(a):
-    i = np.where(a == 0)
+def cross_entropy(a,t):
+    i = np.where(a==0)
     a[i] = 1e-15 # replace 0s so ln doesn't produce infinity
-    e = -np.log(a).sum()
+    e = -t * np.log(a)
     return e
 
 # calculate hidden layer sensitivity
@@ -93,8 +88,8 @@ def learn(weight_old, bias_old, sensitivity, input, learning_rate):
 # split dataset into training, validation, testing sets
 def split(dataset,t,v):
     # split dataset by target class
-    dataset0 = dataset.ix[dataset.ix[:,'target']==0,:]
-    dataset1 = dataset.ix[dataset.ix[:,'target']==1,:]
+    dataset0 = dataset.ix[dataset.ix[:,'window']==0,:]
+    dataset1 = dataset.ix[dataset.ix[:,'window']==1,:]
 
     # build datasets by target class
     train0, validate0, test0 = np.split(dataset0.sample(frac=1),
@@ -122,8 +117,13 @@ def main():
     # import dataset
     cols = ['id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'type']
     glass = pd.read_csv('https://archive.ics.uci.edu/ml/machine-learning-databases/glass/glass.data', names=cols)
-    glass.ix[glass.ix[:, 'type'] <= 4, 'target'] = 0 # assign 0 to window glass
-    glass.ix[glass.ix[:, 'type'] > 4, 'target'] = 1 # assign 1 to non-window glass
+    # create window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] <= 4, 'window'] = 1
+    glass.ix[glass.ix[:, 'type'] > 4, 'window'] = 0
+    # create non-window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] > 4, 'non_window'] = 1
+    glass.ix[glass.ix[:, 'type'] <= 4, 'non_window'] = 0
+    # drop id and class columns
     glass.drop(['id', 'type'], inplace=True, axis=1)
 
[…]

### 2017-06-25 neuron.py
Re-named 'n' in jacobian functions

@@ -42,10 +42,10 @@ class Neuron:
         return a
 
     # softmax derivative jacobian matrix
-    def j_softmax(self,a,n):
-        jacobian = np.empty([n, n])
-        for i in range(n):
-            for j in range(n):
+    def j_softmax(self,a,num):
+        jacobian = np.empty([num, num])
+        for i in range(num):
+            for j in range(num):
                 if i == j:
                     jacobian[i, j] = a[i] * np.sum(a - a[i])
                 else:
@@ -58,8 +58,8 @@ class Neuron:
         return a
 
     # tansig derivative jacobian matrix
-    def j_tansig(self,a,n):
-        jacobian = np.diag(np.ones(n)) - np.diagflat(np.power(a, 2))
+    def j_tansig(self,a,num):
+        jacobian = np.diag(np.ones(num)) - np.diagflat(np.power(a, 2))
         return jacobian
 
 # define cross entropy error calculation function for softmax (assuming targets of 1,0)
@@ -167,10 +167,10 @@ def main():
         cost_t.append(e)
 
         # calculate layer 2 sensitivity
-        s2 = senseo(F_prime=neuron2.j_softmax(a=a2,n=num_neurons2),e=e)
+        s2 = senseo(F_prime=neuron2.j_softmax(a=a2,num=num_neurons2),e=e)
 
         # calculate layer 1 sensitivity
-        s1 = senseh(F_prime=neuron1.j_tansig(a=a1,n=num_neurons1),W_1=W2,s_1=s2)
+        s1 = senseh(F_prime=neuron1.j_tansig(a=a1,num=num_neurons1),W_1=W2,s_1=s2)
 
         # calculate new weight and bias for layer 2
         W2, b2 = learn(weight_old=W2,bias_old=b2,sensitivity=s2,input=a1,learning_rate=alpha)

### 2017-06-25 Paper/paper.tex
Add references to paper

@@
 \documentclass[12pt,halfline,a4paper]{ouparticle}
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[
+backend=biber,
+style=alphabetic,
+sorting=ynt
+]{biblatex}
+\usepackage[nottoc]{tocbibind}
+
+\addbibresource{paper.bib}
 
@@
 \section{Description of the Dataset}
+\label{sec4}
+The dataset was made available for download from the UCI Machine Learning Repository and was created by the USA Forensic Science Service \cite{murphy1994}.
+The purpose of this dataset is to use physicochemical properties to classify whether a certain glass fragment comes from a window or not.
+
 \subsection{Inputs}
+The matrix of inputs contains 214 observations of 9 variables and there is no missing data.
+Of these variables, eight of the nine measure the percent weight that a given elemental oxide makes up of the total glass sample weight.
+All the eight elements except silicon are classified as metals on the periodic table of elements.
+Sodium and potassium are alkali metals whereas magnesium, calcium, and barium alkaline earth metals.
+Aluminum and iron are classified as poor metals and transition metals, respectively.
+The last variable in the input matrix represents the refractive index which measures the speed of light in a transparent medium and is known as Snell's law.
+It can be represented formulaically as the ratio of the velocity of light in a vacuum to the velocity of light in the glass itself: $n = \frac{c}{v}$.
+A more thorough description of target variables is as follows:
+\begin{description}
+\item \textbf{Refractive Index:}
+measures the ratio of the velocity of light in a vacuum to the velocity of light in the glass itself
+\item \textbf{Sodium:}
+represents the percent weight in sodium oxide ($\mathrm{Na_{2}O}$)
+\item \textbf{Magnesium:}
+represents the percent weight in magnesium oxide ($\mathrm{MgO}$)
+\item \textbf{Aluminum:}
+represents the percent weight in aluminum oxide ($\mathrm{Na_{2}O}$)
+\item \textbf{Silicon:}
+represents the percent weight in silicon oxide ($\mathrm{Al_{2}O_{3}}$)
+\item \textbf{Potassium:}
+represents the percent weight in potassium oxide ($\mathrm{K_{2}O}$)
+\item \textbf{Calcium:}
+represents the percent weight in calcium oxide ($\mathrm{Ca_{2}O}$)
+\item \textbf{Barium:}
+represents the percent weight in barium oxide ($\mathrm{Ba_{2}O}$)
+\item \textbf{Iron:}
+represents the percent weight in iron oxide ($\mathrm{Fe_{2}O_{3}}$)
+\end{description}
+
 \subsection{Targets}
+The matrix of targets has 214 observations, one for each observation in the training set, where a given target is denoted by $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$ if the glass sample comes from a window and $\begin{bmatrix} 0 \\ 1 \end{bmatrix}$ otherwise.
+Note that the targets are two-dimensional instead of the more common one-dimensional binary encoding.
+This two-dimensional encoding allows us to have only one neuron firing at a time and tends to result in marginally better performance.
 
@@
 \section{Conclusion}
-
-\subsection{General guidelines}
-\label{sec3.1}
-
-By default, all of the options within \verb+article.cls+ are available
-with this class file. This class file provides the following additional options.
-
-\begin{description}
-
-\item \textbf{endnotes:}
-To make all footnotes to endnotes. You may follow the same
-coding \verb+\footnote{text}+ for both footnotes and endnotes. Once you use this option
-you have to use the \verb+\theendnotes+ command at the place where all the endnotes
-have to be set in your paper.
-
-\item \textbf{numbib:}
-This is the default option that numbers the bibliography items;
-this option does nothing with natbib and other packages.
-
-\end{description}
-
-\subsection{Sections and subsections}
-\label{sec3.6}
-
- If you want to be able to make reference to that section,
-then you need to \texttt{label} it (see Section \ref{sec3.14}).
-
-\subsubsection*{\LaTeX\ text formatting commands}
-\begin{tabular}{ll@{\hskip60pt}ll}
-\verb+\textit+  & Italics      &\verb+\textsf+  & Sans Serif\\
-\verb+\textbf+  & Boldface     &\verb+\textsc+  & Small Caps\\
-\verb+\texttt+  & Typewriter   &\verb+\textmd+  & Medium Series\\
-\verb+\textrm+  & Roman        &\verb+\textnormal+ & Normal Series\\
-\verb+\textsl+  & Slanted      &\verb+\textup+  & Upright Series
-\end{tabular}
-
-\subsubsection*{\LaTeX\ math formatting commands}
-\begin{tabular}{ll@{\qquad}ll}
-\verb+\mathit+     & Math Italics            &\verb+\mathfrak+   & Fraktur\\
-\verb+\mathbf+     & Math Boldface       &\verb+\mathbb+     & Blackboard Bold\\
-\verb+\mathtt+     & Math Typewriter     &\verb+\mathnormal+ & Math Normal\\
-\verb+\mathsf+     & Math Sans Serif     &\verb+\boldsymbol+ & Bold math for Greek letters\\
-\verb+\mathcal+    & Calligraphic        &                   & and other symbols
-\end{tabular}
 
@@
 
-\addcontentsline{toc}{section}{References}
-\begin{thebibliography}{0}
-\bibitem{bib1}
-Goossens, M., F. Mittelbach, and A. Samarin: {\em The {\LaTeX} Companion}.
-Addison-Wesley, Reading, MA, USA, 1994.
-\bibitem{bib2}
-Knuth, D.E: {\em The {\TeX}book}. Addison-Wesley, Reading, MA, USA, 1984.
-\bibitem{bib3}
-Lamport, L.: {\em {\LaTeX} -- A Document Preparation System -- User's
-Guide and Reference Manual}. Addison-Wesley, Reading, MA, USA, 1985.
-\bibitem{bib4}
-Smith, I.N., R.S. Johnes, and W.P. Hines: 1992, `Title of the Article',
-\textit{Journal Title in Italics} \textbf{Vol. no. X}, pp. 00--00
-\end{thebibliography}
+\printbibliography[heading=bibintoc, title={References}]
+
+%\addcontentsline{toc}{section}{References}
+%\begin{thebibliography}{0}
+%\bibitem{bib1}
+%Murphy, P.M., Aha, D.W. (1994): {\em UCI Repository of Machine Learning Databases.} 
+%[http://www.ics.uci.edu/~mlearn/MLRepository.html]. Irvine, CA: University of California, Department of Information and Computer Science.
+%\bibitem{bib2}
+%Knuth, D.E: {\em The {\TeX}book}. Addison-Wesley, Reading, MA, USA, 1984.
+%\bibitem{bib3}
+%Lamport, L.: {\em {\LaTeX} -- A Document Preparation System -- User's
+%Guide and Reference Manual}. Addison-Wesley, Reading, MA, USA, 1985.
+%\bibitem{bib4}
+%Smith, I.N., R.S. Johnes, and W.P. Hines: 1992, `Title of the Article',
+%\textit{Journal Title in Italics} \textbf{Vol. no. X}, pp. 00--00
+%\end{thebibliography}
 

### 2017-06-25 neuron.py
minor tweaks
I’m stopping for the evening. the code currently works, however, I
don’t think I’m implementing the learning rule correctly. I believe
that everything up through calculating the cross entropy is correct,
but now sure i’m doing the weight and bias updates right. Would love
for you guys to look at that.

@@ -70,8 +70,8 @@ def cross_entropy(a,t):
     return e
 
 # calculate hidden layer sensitivity
-def senseh(F_prime,W_1,s_1):
-    s = F_prime * np.transpose(W_1) * s_1
+def senseh(F_prime,W,s):
+    s = F_prime * np.transpose(W) * s
     return s
 
 # calculate output layer sensitivity
@@ -130,10 +130,10 @@ def main():
     train, validate, test = split(glass, t=0.7, v=0.15)
 
     # specify number of neurons in each layer and the learning rate
-    num_neurons1 = 2
+    num_neurons1 = 10
     num_neurons2 = 2
-    alpha = 0.2
-    iterations = 100
+    alpha = 0.05
+    iterations = 25
 
     # initialize weight and bias randomly for each layer from -0.5 to 0.5
     # W1 number of columns matches training set columns, less final two
@@ -142,9 +142,9 @@ def main():
     W2 = np.matrix(np.random.rand(num_neurons2,num_neurons1) - 0.5)
     b2 = np.matrix(np.random.rand(num_neurons2,1) - 0.5)
 
-    # initialize cost lists for training and validation sets
-    cost_t = []
-    cost_v = []
+    # initialize cross entropy lists for training and validation sets
+    ce_t = []
+    ce_v = []
 
     # training & validation
     for epoch in range(iterations):
@@ -164,28 +164,32 @@ def main():
         # calculate error of each iteration and update cost total
         target = np.matrix([train.iloc[epoch,-2:]])
         e = cross_entropy(a2,target)
-        cost_t.append(e)
+        ce_t.append(np.asscalar(e))
+        e = np.concatenate([e,e])
 
         # calculate layer 2 sensitivity
         s2 = senseo(F_prime=neuron2.j_softmax(a=a2,num=num_neurons2),e=e)
 
         # calculate layer 1 sensitivity
-        s1 = senseh(F_prime=neuron1.j_tansig(a=a1,num=num_neurons1),W_1=W2,s_1=s2)
+        s1 = senseh(F_prime=neuron1.j_tansig(a=a1,num=num_neurons1),W=W2,s=s2)
 
         # calculate new weight and bias for layer 2
         W2, b2 = learn(weight_old=W2,bias_old=b2,sensitivity=s2,input=a1,learning_rate=alpha)
[…]

### 2017-06-25 neuron.py
Normalization
I added lines to normalize the input data.

@@ -85,8 +85,12 @@ def learn(weight_old, bias_old, sensitivity, input, learning_rate):
     bias_new = bias_old - learning_rate * sensitivity
     return  weight_new, bias_new
 
-# split dataset into training, validation, testing sets
+# normailize data and split dataset into training, validation, testing sets
 def split(dataset,t,v):
+    # normalize dataset into range between -1 and 1
+    norm = dataset.ix[:, :-2].apply(lambda x: -1 + 2 * (x - x.min()) / (x.max() - x.min()), axis=0)
+    dataset = pd.concat([norm,dataset.ix[:,-2:]],axis=1)
+
     # split dataset by target class
     dataset0 = dataset.ix[dataset.ix[:,'window']==0,:]
     dataset1 = dataset.ix[dataset.ix[:,'window']==1,:]

### 2017-06-25 Paper/paper.tex
Add network figure
Update paper

@@
 \section{Description of the Network Architecture and Training Algorithm}
+Once the data is preprocessed, the next step is to decide on or create a network architecture.
+The basic network architecture is determined by the problem we wish to solve.
+Once the basic network architecture is determined, we decide how many layers, how many neurons in each layer, how many outputs the network should have, and what kind of performance index function we should use for training \cite{hagan2014}.
 
 \subsection{Network Architecture}
-\subsection{Training Algorith}
+The standard neural network architecture for pattern recognition problems is the multi-layer perceptron with tangent-sigmoid transfer functions in the hidden layers and \emph{softmax} transfer functions in the output layer.
+For most problems, including a fairly simple one like ours, one hidden layer usually suffices.
+Thus, we will implement a two-layer perceptron.
+If the results of the network are unsatisfactory after training and testing with one hidden layer, we will retrain with an additional hidden layer, but we do not anticipate having to do this.
+The \emph{tansig} transfer function is usually preferred to the \emph{logsig} transfer function in the hidden layers since it produces outputs (which are inputs to the next layer) that are centered near zero, whereas the \emph{logsig} transfer function always produces positive outputs.
+
+We also need to select the number of neurons in each layer.
+The number of neurons we use in the output layer should be the same as the size of the target vector.
+In our case, this means we should use two neurons in the output layer.
+On the other hand, the number of neurons we use in the hidden layer is directly proportional to the the complexity of the decision boundary being implemented.
+Since we do not know the complexity of the decision boundary needed to classify these glass samples before training, we begin with ten neurons, which may be more than we need, and leverage early stopping techniques to prevent overfitting \cite{hagan2014}.
+
+Now that we have chosen a network architecture, we can calculate the network output. The output from the hidden layer (the input to the output layer) can be calculated as 
+\[
+\mathbf{a^{1}} = \textbf{logsig}(\mathbf{W^{1}}\mathbf{p} + \mathbf{b^{1}}), 
+\] 
+while the output from the output layer is 
+\[ 
+\mathbf{a^{2}} = \textbf{logsig}(\mathbf{W^{2}}\mathbf{a^{1}} + \mathbf{b^{2}}). 
+\]
+The network architecture can be seen in \emph{Figure 1} below. 
+\begin{figure}[H]
+\includegraphics[width = 6in]{figs/network.png}
+\caption{Two-Layer Perceptron Used to Classify Glass Samples}
+\end{figure}
+
+In the 1950's, Frank Rosenblatt developed the perceptron learning rule for training networks to solve pattern recognition problems.
+While his developments were monumental at the time, we only implement the perceptron learning rule when the input vectors are \emph{linearly separable}.
+Widrow and Hoff's learning rule suffered from the same disadvantage.
+When we have more complicated decision boundaries we are trying to implement and have multiple layers, a more robust learning rule ought to be used.
+This problem prompted Paul Werbos to derive the backpropagation algorithm in his thesis in the 1970's.
+Since then, multilayer perceptrons trained by the backpropagation algorithm have become the most popular neural network.
+\cite{hagan2014}.
+
+\subsection{Training Algorithm}
+We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
+For multi-layer networks, the Levenberg-Marquardt algorithm is often used, but it does not work well for pattern recognition as the transfer function in the output layer is operating outside the linear region.
+The scaled conjugate gradient algorithm is a special type of backpropagation.
+In backpropagation, we adjust the weights in the direction of steepest descent
 
@@
 
-\subsection{Figures and tables}
+\section{Figures and tables}
 \label{sec3.9}
@@
 
-%\addcontentsline{toc}{section}{References}
-%\begin{thebibliography}{0}
-%\bibitem{bib1}
-%Murphy, P.M., Aha, D.W. (1994): {\em UCI Repository of Machine Learning Databases.} 
-%[http://www.ics.uci.edu/~mlearn/MLRepository.html]. Irvine, CA: University of California, Department of Information and Computer Science.
-%\bibitem{bib2}
-%Knuth, D.E: {\em The {\TeX}book}. Addison-Wesley, Reading, MA, USA, 1984.
-%\bibitem{bib3}
-%Lamport, L.: {\em {\LaTeX} -- A Document Preparation System -- User's
-%Guide and Reference Manual}. Addison-Wesley, Reading, MA, USA, 1985.
-%\bibitem{bib4}
-%Smith, I.N., R.S. Johnes, and W.P. Hines: 1992, `Title of the Article',
-%\textit{Journal Title in Italics} \textbf{Vol. no. X}, pp. 00--00
-%\end{thebibliography}
-
 \pagebreak

### 2017-06-25 Paper/paper.tex
Typed up Literature Review

@@
 \subsection{Literature Review}
-\label{sec3}
+\label{sec2.1}
+The dataset we will use for this analysis was initially employed in Evett and Spiehler's paper \emph{Rule Induction in Forensic Science} (1987) \cite{evett1987}.
+They recognized the usefulness to a forensic crime lab of classifying glass fragments based on refractive index and chemical composition.
+This would, for example, allow the lab to ascertain whether samples gathered on a suspect?s clothing came from a window, potentially indicating they had broken it, or from another source, like a broken bottle.
+
+In their experiment, Evett and Spiehler wished to see if the Bionic Evolutionary Algorithm Generating Logical Expressions (BEAGLE) machine learning algorithm could correctly classify the glass---first as either window or non-window, and then into a second level of sub-categories.
+Our project focuses on the former and leaves the latter as a future exercise.
+The BEAGLE algorithm Evett and Spiehler used to train the network uses a series of logical ``and'' statements to chain rules together based on the inputs.
+They offer up the following example of a rule:
+\[ 
+\{(\mathrm{Fe} \leq \mathrm{Na}) \text{ and } [\mathrm{K} > (\mathrm{Fe} \cdot 650)]\}, 
+\]
+where $\mathrm{Fe}$, $\mathrm{Na}$, and $\mathrm{K}$ are the percent composition of iron, sodium, and potassium, respectively, of each glass sample.
+Evett and Spiehler found that the BEAGLE algorithm outperformed the $k$-Nearest Neighbors algorithm (with $ k = 3$) and Linear Discriminant Analysis (LDA)---the models they used for baseline performance measures.
+
+The Department of Justice regularly issues research grants for the elemental analysis of glass.
+In 2012, they issued one such grant to researchers at Florida International University (FIU) to work with the Miami-Dade County Police Department \cite{almirall2012}.
+The researchers note that analysis of small quantities of materials has become an important yet underutilized type of evidence at many crime scenes including hit-and-run accidents and other violent crimes.
+The ability to classify different types of glass could be of vital importance in the case of a hit-and-run.
+Further, the group of researchers attempted to compare the discrimination power between the methods used in most forensic laboratories for glass analysis.
+Their aim was to create a more ``standard'' method that can be used by the operational forensic laboratory and a ``match criteria'' for use in routine casework situations.
+
+Maureen Bottrell, a geologist and forensic scientist at the FBI Laboratory released a report in 2009 documenting the background information that ought to be used when comparing glass samples with data \cite{bottrell2009}.
+She notes that the vast majority of raw materials used to make glass are derived geologically and that North American glass makers use more than 20 million tons of raw materials annually.
+All of these materials contain several impurities that result in perceived differences in glass products.
+
+Bottrell writes that physical properties such as color, curvature, fluorescence, thickness, and surface features should first be used to determine if the material fragments are glass.
+Once we know a sample is glass, Bottrell recommends using optical properties, particle immersion, density, and elemental analysis to differentiate between types of glass.
+In this study, we focus on optical properties, specifically the refractive index, and elemental analysis to classify whether glass samples came from a window or not.\footnote{See section \ref{sec3.1} for more on the refractive index.}
+
+Since glass is made up of several raw materials and certain elements impart specific properties, we can find out a lot about the glass if we analyze the chemical composition.
+Glass made on the same manufacturing line over a period of time can often have highly variable properties as mixtures of raw materials can have drastically different chemical compositions.
 
 \section{Description of the Dataset}
-\label{sec4}
+\label{sec3}
 The dataset was made available for download from the UCI Machine Learning Repository and was created by the USA Forensic Science Service \cite{murphy1994}.
@@
 \subsection{Inputs}
+\label{sec3.1}
 The matrix of inputs contains 214 observations of 9 variables and there is no missing data.
@@
 \subsection{Targets}
+\label{sec3.2}
 The matrix of targets has 214 observations, one for each observation in the training set, where a given target is denoted by $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$ if the glass sample comes from a window and $\begin{bmatrix} 0 \\ 1 \end{bmatrix}$ otherwise.
@@
 \section{Description of the Network Architecture and Training Algorithm}
+\label{sec4}
 Once the data is preprocessed, the next step is to decide on or create a network architecture.
@@
 \subsection{Network Architecture}
+\label{sec4.1}
 The standard neural network architecture for pattern recognition problems is the multi-layer perceptron with tangent-sigmoid transfer functions in the hidden layers and \emph{softmax} transfer functions in the output layer.
@@
 \[ 
-\mathbf{a^{2}} = \textbf{logsig}(\mathbf{W^{2}}\mathbf{a^{1}} + \mathbf{b^{2}}). 
+\mathbf{a^{2}} = \textbf{softmax}(\mathbf{W^{2}}\mathbf{a^{1}} + \mathbf{b^{2}}). 
 \]
-The network architecture can be seen in \emph{Figure 1} below. 
+The network architecture can be seen in \textbf{Figure 1} below. 
 \begin{figure}[H]
@@
 
-In the 1950's, Frank Rosenblatt developed the perceptron learning rule for training networks to solve pattern recognition problems.
-While his developments were monumental at the time, we only implement the perceptron learning rule when the input vectors are \emph{linearly separable}.
-Widrow and Hoff's learning rule suffered from the same disadvantage.
+\subsubsection{History of the Perceptron}
+\label{sec4.1.1}
+Before discussing the training algorithm, we offer up a brief history of the perceptron.
+Much of the modern interpretation of neural networks is credited in large part to Warren McCulloch and Walter Pitts who showed that networks of artificial neurons could calculate any function.
+We still use the fundamental feature of their model in which a weighted sum of inputs is compared to a threshold to determine the output of a neuron \cite{hagan2014}.
+
+One of the first applications of neural networks came in the 1950's when Frank Rosenblatt developed the perceptron network and the corresponding learning rule to solve pattern recognition problems.
+While his developments were monumental at the time, several researchers showed that a single-layer perceptron and the learning rule could not solve certain problems.
+Specifically, the error will never converge to zero when using the perceptron learning rule if the input vectors are \emph{linearly inseparable}.
+It wasn't until the 1980's that multi-layer perceptron networks and more complex learning rules were proposed that could solve these problems.
+
+%Widrow and Hoff's learning rule suffered from the same disadvantage.
 When we have more complicated decision boundaries we are trying to implement and have multiple layers, a more robust learning rule ought to be used.
-This problem prompted Paul Werbos to derive the backpropagation algorithm in his thesis in the 1970's.
 Since then, multilayer perceptrons trained by the backpropagation algorithm have become the most popular neural network.
@@
 \subsection{Training Algorithm}
-We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
+\label{sec4.2}
+%We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
 For multi-layer networks, the Levenberg-Marquardt algorithm is often used, but it does not work well for pattern recognition as the transfer function in the output layer is operating outside the linear region.
 The scaled conjugate gradient algorithm is a special type of backpropagation.
-In backpropagation, we adjust the weights in the direction of steepest descent
+In backpropagation, we adjust the weights in the direction of steepest descent...
 
 \section{Experimental Setup}
+\label{sec5}
 
 \subsection{Data Preprocessing}
+\label{sec5.1}
+
 \subsection{Implementation of the Network}
+\label{sec5.2}
+
 \subsection{Performance Index}
+\label{sec5.3}
 
 \section{Results}
+\label{sec6}
 
 \section{Conclusion}
-
-\section{Figures and tables}
-\label{sec3.9}
-
-The following is an example of typesetting a table.
-
-\begin{verbatim}
-\begin{table}
-\caption{Table caption text.}
-\label{key}
-The table matter goes here.
-\end{table}
-\end{verbatim}
-
-As always with \LaTeX, the \verb+\label+ must be after the
-\verb+\caption+, and inside the figure or table environment. The reference for
-figures and tables inside text can be made using the \verb|\ref{key}| command.
-
-\subsection{Equations}\label{sec3.10}
-
-For example, if you type
-\begin{verbatim}
-\begin{equation}\label{eq1}
-\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = [\sigma r_2/(2\mu_0)]
-\int^{\infty}_0\exp(-\lambda|z_j-z_i|)\lambda^{-1}J_1 (\lambda r_2)J_0
-(\lambda r_i\,\lambda {\rm d}\lambda)
-\end{equation}
-\end{verbatim}
-then you will get the following output:
-\begin{equation}\label{eq1}
-\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = [\sigma r_2/(2\mu_0)]\int^{\infty}_0
-\exp(-\lambda|z_j-z_i|)\lambda^{-1}J_1 (\lambda r_2)J_0 (\lambda r_i\,\lambda {\rm d}\lambda)
-\end{equation}
-
-\subsection{Listings}
-\label{sec3.12}
-
-Another frequently displayed structure is a list. The
-following is an example of an \emph{itemized} list.
-\begin{itemize}
-   \item This is the first item of an itemized list.
-         Each item in the list is marked with a
-         `$\bullet$'.
-
-   \item This is the second item of the list. It
-         contains another list nested inside it. The
-         inner list is an \emph{enumerated} list.
-         \begin{enumerate}
-            \item This is the first item of an enumerated
-                  list that is nested within the
-                  itemized list.
-
-            \item This is the second item of the inner list.
-                  \LaTeX\ allows you to nest lists
-                  deeper than you really should.
-         \end{enumerate}
-         This is the rest of the second item of the
-         outer list. It is no more interesting than
-         any other part of the item.
-   \item This is the third item of the list.
-\end{itemize}
+\label{sec7}
 

### 2017-06-25 MATLAB docs/trainscg.m
Scaled Conjugate Gradient Function

@@ -1,0 +1,315 @@
+function [out1,out2] = trainscg(varargin)
+%TRAINSCG Scaled conjugate gradient backpropagation.
+%
+%  <a href="matlab:doc trainscg">trainscg</a> is a network training function that updates weight and
+%  bias values according to the scaled conjugate gradient method.
+%
+%  [NET,TR] = <a href="matlab:doc trainscg">trainscg</a>(NET,X,T) takes a network NET, input data X
+%  and target data T and returns the network after training it, and a
+%  a training record TR.
+%
+%  [NET,TR] = <a href="matlab:doc trainscg">trainscg</a>(NET,X,T,Xi,Ai,EW) takes additional optional
+%  arguments suitable for training dynamic networks and training with
+%  error weights.  Xi and Ai are the initial input and layer delays states
+%  respectively and EW defines error weights used to indicate
+%  the relative importance of each target value.
+%
+%  Training occurs according to training parameters, with default values.
+%  Any or all of these can be overridden with parameter name/value argument
+%  pairs appended to the input argument list, or by appending a structure
+%  argument with fields having one or more of these names.
+%    epochs            1000  Maximum number of epochs to train
+%    show                25  Epochs between displays
+%    showCommandLine  false  Generate command-line output
+%    showWindow        true  Show training GUI
+%    goal                 0  Performance goal
+%    time               inf  Maximum time to train in seconds
+%    min_grad          1e-6  Minimum performance gradient
+%    max_fail             6  Maximum validation failures
+%    sigma           5.0e-5  Determines change in weight for second derivative approximation.
+%    lambda          5.0e-7  Parameter for regulating the indefiniteness of the Hessian.
+%
+%  To make this the default training function for a network, and view
+%  and/or change parameter settings, use these two properties:
+%
+%    net.<a href="matlab:doc nnproperty.net_trainFcn">trainFcn</a> = 'trainscg';
+%    net.<a href="matlab:doc nnproperty.net_trainParam">trainParam</a>
+%
+%  See also TRAINGDM, TRAINGDA, TRAINGDX, TRAINLM, TRAINRP,
[…]

### 2017-06-25 neuron.py
Confusion matrix
Added testing and confusion matrix.

@@ -62,11 +62,18 @@ class Neuron:
         jacobian = np.diag(np.ones(num)) - np.diagflat(np.power(a, 2))
         return jacobian
 
+    # classify output
+    def classify(self,a):
+       c = np.where(a[0]>=0.5,1,0)
+       c = np.asscalar(c)
+       return c
+
 # define cross entropy error calculation function for softmax (assuming targets of 1,0)
 def cross_entropy(a,t):
     i = np.where(a==0)
     a[i] = 1e-15 # replace 0s so ln doesn't produce infinity
     e = -t * np.log(a)
+    #e = - np.multiply(np.transpose(t), np.log(a))
     return e
 
 # calculate hidden layer sensitivity
@@ -136,8 +143,8 @@ def main():
     # specify number of neurons in each layer and the learning rate
     num_neurons1 = 10
     num_neurons2 = 2
-    alpha = 0.05
-    iterations = 25
+    alpha = 0.1
+    iterations = 100
 
     # initialize weight and bias randomly for each layer from -0.5 to 0.5
     # W1 number of columns matches training set columns, less final two
@@ -149,13 +156,19 @@ def main():
     # initialize cross entropy lists for training and validation sets
     ce_t = []
     ce_v = []
+    # initialize validation set indexer
+    index_v = 0
 
     # training & validation
     for epoch in range(iterations):
 
-        # create input matrix from training dataset with number of columns to match neurons
-        input = np.matrix(train.iloc[epoch,:-2])
-        #input = np.concatenate([input]*num_neurons1)
+        # if number of iterations is longer than dataset, loop back to the beginning
+        index = epoch
+        if index >= len(train):
+            index = epoch - len(train)
+
+        # create input matrix from training dataset
+        input = np.matrix(train.iloc[index,:-2])
         input = np.transpose(input)
 
         # first layer
@@ -166,7 +179,7 @@ def main():
         a2 = neuron2.softmax()
 
         # calculate error of each iteration and update cost total
-        target = np.matrix([train.iloc[epoch,-2:]])
+        target = np.matrix(train.iloc[index,-2:])
         e = cross_entropy(a2,target)
         ce_t.append(np.asscalar(e))
         e = np.concatenate([e,e])
@@ -182,18 +195,57 @@ def main():
         # calculate new weight and bias for layer 1
[…]

### 2017-06-25 neuron.py
Plot labels

@@ -219,11 +219,12 @@ def main():
 
 
     fig, ax = plt.subplots(ncols=1, nrows=1, figsize=[8, 8])
-    ax.plot(np.arange(0,iterations),ce_t)
-    ax.plot(np.arange(0,len(ce_v)),ce_v,'k.')
+    ax.plot(np.arange(0,iterations),ce_t,label='Training')
+    ax.plot(np.arange(0,len(ce_v)),ce_v,'k.',label='Validation')
     plt.xlabel('epochs')
     plt.ylabel('cross entropy')
     plt.grid()
+    ax.legend()
     plt.show()
 
     # initialize confusion matrix series

### 2017-06-26 neuron.py
Batch update & early stopping
I re-configured the process to calculate the average gradient after a
full pass through the dataset before updating. I also implemented early
stopping, though with a minimum of 5 iterations.

@@ -38,7 +38,8 @@ class Neuron:
 
     # softmax transfer function a = e^n/sum(e^n)
     def softmax(self):
-        a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
+        #a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
+        a = np.exp(self.net_input()) / np.exp(self.net_input()).sum(axis=0)
         return a
 
     # softmax derivative jacobian matrix
@@ -65,7 +66,6 @@ class Neuron:
     # classify output
     def classify(self,a):
        c = np.where(a[0]>=0.5,1,0)
-       c = np.asscalar(c)
        return c
 
 # define cross entropy error calculation function for softmax (assuming targets of 1,0)
@@ -73,7 +73,6 @@ def cross_entropy(a,t):
     i = np.where(a==0)
     a[i] = 1e-15 # replace 0s so ln doesn't produce infinity
     e = -t * np.log(a)
-    #e = - np.multiply(np.transpose(t), np.log(a))
     return e
 
 # calculate hidden layer sensitivity
@@ -127,7 +126,7 @@ def main():
 
     # import dataset
     cols = ['id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'type']
-    glass = pd.read_csv('https://archive.ics.uci.edu/ml/machine-learning-databases/glass/glass.data', names=cols)
+    glass = pd.read_csv('glass.txt', names=cols)
     # create window target column (1 for true; 0 for false)
     glass.ix[glass.ix[:, 'type'] <= 4, 'window'] = 1
     glass.ix[glass.ix[:, 'type'] > 4, 'window'] = 0
@@ -144,7 +143,7 @@ def main():
     num_neurons1 = 10
     num_neurons2 = 2
     alpha = 0.1
-    iterations = 100
+    epoch = 30
 
     # initialize weight and bias randomly for each layer from -0.5 to 0.5
     # W1 number of columns matches training set columns, less final two
@@ -160,93 +159,104 @@ def main():
     index_v = 0
 
     # training & validation
-    for epoch in range(iterations):
+    for j in range(epoch):
 
-        # if number of iterations is longer than dataset, loop back to the beginning
-        index = epoch
-        if index >= len(train):
-            index = epoch - len(train)
+        for i in range(len(train)):
+            # create input matrix from training dataset
+            input = np.matrix(train.iloc[i,:-2])
[…]

### 2017-06-26 Paper/paper.tex
Paper Edits
Paper edits

@@
 If the results of the network are unsatisfactory after training and testing with one hidden layer, we will retrain with an additional hidden layer, but we do not anticipate having to do this.
-The \emph{tansig} transfer function is usually preferred to the \emph{logsig} transfer function in the hidden layers since it produces outputs (which are inputs to the next layer) that are centered near zero, whereas the \emph{logsig} transfer function always produces positive outputs.
+The \emph{tansig} transfer function is usually preferred to the \emph{tansig} transfer function in the hidden layers since it produces outputs (which are inputs to the next layer) that are centered near zero, whereas the \emph{tansig} transfer function always produces positive outputs.
 
@@
 \[
-\mathbf{a^{1}} = \textbf{logsig}(\mathbf{W^{1}}\mathbf{p} + \mathbf{b^{1}}), 
+\mathbf{a^{1}} = \textbf{tansig}(\mathbf{W^{1}}\mathbf{p} + \mathbf{b^{1}}), 
 \] 

### 2017-06-27 neuronLayer.py
New Neutron Layer Routines
still in development

@@ -1,0 +1,279 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+
+# create neuron class
+class NeuronLayer:
+    # define attributes
+    def __init__(self, r, s, f, j):
+        self.r = r # Number of inputs
+        self.s = s # number of neurons
+        self.f = f # Transfer function
+        self.j = j # derivatve transfer function
+
+    def setWeightBias(self, w, b):
+        self.w = w  # set weights
+        self.b = b # set biases
+        self.r = w.shape[1]
+        self.s = w.shape[0]
+        return
+
+    def processInput(self, p):
+        self.p = p
+        self.n = self.w * self.p + self.b
+        self.a = self.f(self.n)
+        return
+
+
+# hardlimit transfer function a = hardlim(n)
+def hardlim(n):
+    if n >= 0:
+        a = 1
+    else:
+        a = 0
+    return a
+
+
+# log-sigmoid transfer function a = logsig(n)
+def logsig(n):
+    return 1 / (1 + np.exp(-n))
+
+
+# linear transfer function
+def purelin(n):
+    return n
+
+# hyperbolic tangent sigmoid transfer function
+def tansig(n):
+    return (2 / (1 + np.exp(-2 * n)) - 1)
+
+
+# softmax transfer function a = e^n/sum(e^n)
+def softmax(n):
+    # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
+    return np.exp(n / np.exp(n).sum(axis=0))
+
+# classify output
+def classify(a):
+    c = np.where(a[0] >= 0.5, 1, 0)
+    return c
+
+
+# tansig derivative jacobian matrix
+def j_tansig(a):
+    jacobian = np.diag(np.ones(a.shape[0])) - np.diagflat(np.power(a, 2))
+    return jacobian
+
+
+# softmax derivative jacobian matrix
+def j_softmax(a):
+    jacobian = np.empty([a.shape[0], a.shape[0]])
+    for i in range(jacobian.shape[0]):
+        for j in range(jacobian.shape[1]):
+            if i == j:
+                jacobian[i, j] = a[i]*np.sum(a-a[i])
+            else:
+                jacobian[i, j] = -a[i] * a[j]
+    return jacobian
+
+
+# define cross entropy error calculation function for softmax (assuming targets of 1,0)
+def cross_entropy(a, t):
+    i = np.where(a == 0)
+    a[i] = 1e-15  # replace 0s so ln doesn't produce infinity
[…]

### 2017-06-27 neuronLayer.py
Re-factored Object NeuronLayer still Not Working

@@ -110,7 +110,7 @@ def split(dataset, num_inputs, t, v):
     # normalize dataset into range between -1 and 1
 
     norm = dataset.ix[:, slice(0,num_inputs)].apply(lambda x: -1 + 2 * (x - x.min()) / (x.max() - x.min()), axis=0)
-    dataset = pd.concat([norm, dataset.ix[:,'window']], axis=1)
+    dataset = pd.concat([norm, dataset.ix[:,'window':]], axis=1)
 
     # split dataset by target class
     dataset0 = dataset.ix[dataset.ix[:, 'window'] == 0, :]
@@ -161,8 +161,8 @@ def main():
     num_neurons2 = 2
     alpha = 0.1
     epoch = 30
-    neuron1 = NeuronLayer(num_neurons1, num_inputs , tansig, j_tansig)
-    neuron2 = NeuronLayer(num_neurons2, num_neurons1, softmax, j_softmax)
+    nlayer1 = NeuronLayer(num_neurons1, num_inputs , tansig, j_tansig)
+    nlayer2 = NeuronLayer(num_neurons2, num_neurons1, softmax, j_softmax)
 
     # initialize weight and bias randomly for each layer from -0.5 to 0.5
     # W1 number of columns matches training set columns, less final two
@@ -183,20 +183,20 @@ def main():
         for i in range(len(train)):
 
             #set weights and biases
-            neuron1.setWeightBias(w=W1, b=b1)
-            neuron2.setWeightBias(w=W2, b=b2)
+            nlayer1.setWeightBias(w=W1, b=b1)
+            nlayer2.setWeightBias(w=W2, b=b2)
 
             # create input matrix from training dataset
             input = np.matrix(train.iloc[i, slice(0,num_inputs)]).transpose()
 
             # process network
-            neuron1.processInput(input)
-            neuron2.processInput(neuron1.a)
+            nlayer1.processInput(input)
+            nlayer2.processInput(nlayer1.a)
 
 
             # calculate error of each iteration and update cost total
-            target = np.matrix(train.iloc[i, slice(0,num_inputs)])
-            e = cross_entropy(neuron2.a, target)
+            target = np.matrix(train.iloc[i, -2:])
+            e = cross_entropy(nlayer2.a, target)
             if i == 0:
                 e_all = e
             else:
@@ -204,14 +204,14 @@ def main():
             e = np.concatenate([e, e])
 
[…]

### 2017-06-27 DataPrep.py
Refactored the Code Same Errors

@@ -1,0 +1,16 @@
+import pandas as pd
+import numpy as np
+
+def GlassImport():
+    # import dataset
+    cols = ['id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'type']
+    glass = pd.read_csv('glass.txt', names=cols)
+    # create window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] <= 4, 'window'] = 1
+    glass.ix[glass.ix[:, 'type'] > 4, 'window'] = 0
+    # create non-window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] > 4, 'non_window'] = 1
+    glass.ix[glass.ix[:, 'type'] <= 4, 'non_window'] = 0
+    # drop id and class columns
+    glass.drop(['id', 'type'], inplace=True, axis=1)
+    return glass

### 2017-06-27 neuronLayer.py

@@ -1,169 +1,27 @@
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
-
-
-# create neuron class
-class NeuronLayer:
-    # define attributes
-    def __init__(self, r, s, f, j):
-        self.r = r # Number of inputs
-        self.s = s # number of neurons
-        self.f = f # Transfer function
-        self.j = j # derivatve transfer function
-
-    def setWeightBias(self, w, b):
-        self.w = w  # set weights
-        self.b = b # set biases
-        self.r = w.shape[1]
-        self.s = w.shape[0]
-        return
-
-    def processInput(self, p):
-        self.p = p
-        self.n = self.w * self.p + self.b
-        self.a = self.f(self.n)
-        return
-
-
-# hardlimit transfer function a = hardlim(n)
-def hardlim(n):
-    if n >= 0:
-        a = 1
-    else:
-        a = 0
-    return a
-
-
-# log-sigmoid transfer function a = logsig(n)
-def logsig(n):
-    return 1 / (1 + np.exp(-n))
-
-
-# linear transfer function
-def purelin(n):
-    return n
-
-# hyperbolic tangent sigmoid transfer function
-def tansig(n):
-    return (2 / (1 + np.exp(-2 * n)) - 1)
-
-
-# softmax transfer function a = e^n/sum(e^n)
-def softmax(n):
-    # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-    return np.exp(n / np.exp(n).sum(axis=0))
-
-# classify output
-def classify(a):
-    c = np.where(a[0] >= 0.5, 1, 0)
-    return c
-
-
-# tansig derivative jacobian matrix
-def j_tansig(a):
-    jacobian = np.diag(np.ones(a.shape[0])) - np.diagflat(np.power(a, 2))
-    return jacobian
-
-
-# softmax derivative jacobian matrix
-def j_softmax(a):
-    jacobian = np.empty([a.shape[0], a.shape[0]])
-    for i in range(jacobian.shape[0]):
-        for j in range(jacobian.shape[1]):
-            if i == j:
-                jacobian[i, j] = a[i]*np.sum(a-a[i])
-            else:
-                jacobian[i, j] = -a[i] * a[j]
-    return jacobian
-
-
-# define cross entropy error calculation function for softmax (assuming targets of 1,0)
-def cross_entropy(a, t):
-    i = np.where(a == 0)
[…]

### 2017-06-28 Python Code/NLObjects.py
Organize

@@ -52,7 +52,7 @@ def tansig(n):
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
     # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-    return np.exp(n / np.exp(n).sum(axis=0))
+    return np.exp(n) / np.exp(n).sum(axis=0)
 
 # classify output
 def classify(a):

### 2017-06-28 Python Code/neuronLayer.py

@@ -5,12 +5,13 @@ import NLObjects as nl
 import DataPrep as dp
 import Split as sp
 
-# normailize data and split dataset into training, validation, testing sets
+# normalize data and split dataset into training, validation, testing sets
+
 
 def main():
     # import dataset
     glass = dp.GlassImport()
-    num_inputs = glass.shape[1] -2
+    num_inputs = glass.shape[1] - 2
     # split into training, validate, testing sets
     train, validate, test = sp.split(glass, num_inputs, t=0.7, v=0.15)
 
@@ -19,7 +20,7 @@ def main():
     num_neurons2 = 2
     alpha = 0.1
     epoch = 30
-    nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs , nl.tansig, nl.j_tansig)
+    nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)
     nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)
     np.random.seed(0)
     # initialize weight and bias randomly for each layer from -0.5 to 0.5

### 2017-06-28 Python Code/NLObjects.py
Fix transfer functions and gradients

@@ -46,33 +46,37 @@ def purelin(n):
 
 # hyperbolic tangent sigmoid transfer function
 def tansig(n):
-    return (2 / (1 + np.exp(-2 * n)) - 1)
-
+    e_pos = np.exp(n)
+    e_neg = np.exp(-n)
+    return (e_pos - e_neg) / (e_pos + e_neg)
 
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
     # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-    return np.exp(n) / np.exp(n).sum(axis=0)
+    return np.exp(n) / np.sum(np.exp(n))
 
 # classify output
 def classify(a):
     c = np.where(a[0] >= 0.5, 1, 0)
     return c
 
-
 # tansig derivative jacobian matrix
 def j_tansig(a):
-    jacobian = np.diag(np.ones(a.shape[0])) - np.diagflat(np.power(a, 2))
-    return jacobian
-
+    tg_diags = 1 - np.tanh(a) ** 2
+    tg = np.zeros((a.shape[0], a.shape[0]))
+    np.fill_diagonal(tg, tg_diags)
+    return tg
 
 # softmax derivative jacobian matrix
 def j_softmax(a):
-    jacobian = np.empty([a.shape[0], a.shape[0]])
+    jacobian = np.zeros((a.shape[0], a.shape[0]))
     for i in range(jacobian.shape[0]):
-        for j in range(jacobian.shape[1]):
+        for j in range(jacobian.shape[0]):
             if i == j:
-                jacobian[i, j] = a[i]*np.sum(a-a[i])
+                x = np.zeros((a.shape[0]))
+                for k in range(jacobian.shape[0]):
+                    x[k] = a[k] - a[i]
+                jacobian[i,j] = a[i] * np.sum(x)
             else:
                 jacobian[i, j] = -a[i] * a[j]
     return jacobian

### 2017-06-28 Python Code/NLObjects.py
Fixed transit jacobson function

@@ -62,7 +62,7 @@ def classify(a):
 
 # tansig derivative jacobian matrix
 def j_tansig(a):
-    tg_diags = 1 - np.tanh(a) ** 2
+    tg_diags = 1 - np.power(np.tanh(a),2)
     tg = np.zeros((a.shape[0], a.shape[0]))
     np.fill_diagonal(tg, tg_diags)
     return tg

### 2017-06-28 Python Code/neuron.py

@@ -215,6 +215,7 @@ def main():
         neuron2 = Neuron(input=a1, weight=W2, bias=b2)
         a2 = neuron2.softmax()
         target = np.matrix(validate.iloc[:, -2:])
+        e = cross_entropy(a2,target)
         # compute errors
         for i in range(len(target)):
             e = cross_entropy(a2[:,i],target[i])

### 2017-06-28 Paper/paper.tex
edits

@@
 
-\section{Introduction}
+\section{Literature Review}
 \label{sec2}
-An overview of the project and an outline of the report (I like to write intro after I finish a project).
-
-\subsection{Literature Review}
-\label{sec2.1}
 The dataset we will use for this analysis was initially employed in Evett and Spiehler's paper \emph{Rule Induction in Forensic Science} (1987) \cite{evett1987}.

### 2017-06-28 Python Code/NLObjects.py
Fixed classification and validation

@@ -53,11 +53,11 @@ def tansig(n):
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
     # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-    return np.exp(n) / np.sum(np.exp(n))
+    return np.exp(n) / np.sum(np.exp(n), axis=0)
 
 # classify output
 def classify(a):
-    c = np.where(a[0] >= 0.5, 1, 0)
+    c = np.where(a[0] >= 0.5, 0, 1)
     return c
 
 # tansig derivative jacobian matrix
@@ -87,6 +87,7 @@ def cross_entropy(a, t):
     i = np.where(a == 0)
     a[i] = 1e-15  # replace 0s so ln doesn't produce infinity
     e = -t * np.log(a)
+    #e = np.matrix([[-t[0,0] * np.log(a[0,0])], [-t[0,1] * np.log(a[1,0])]])
     return e
 
 

### 2017-06-28 Python Code/neuronLayer.py

@@ -63,6 +63,7 @@ def main():
             else:
                 e_all = np.concatenate((e_all, e), axis=1)
             e = np.concatenate([e, e])
+            #e = np.asscalar(e)
 
             # calculate layer 2 sensitivity
             s2 = nl.senseo(F_prime=nlayer2.j(nlayer2.a), e=e)

### 2017-06-28 Python Code/neuronLayer.py
Latest update

@@ -33,8 +33,6 @@ def main():
     # initialize cross entropy lists for training and validation sets
     ce_t = []
     ce_v = []
-    # initialize validation set indexer
-    index_v = 0
 
     # training & validation
     nlayer1.setWeightBias(w=W1, b=b1)

### 2017-06-29 Python Code/NLObjects.py
Code updates

@@ -15,14 +15,14 @@ class NeuronLayer:
         self.s = w.shape[0]
         return
 
-    def processInput(self, p):
+    def FP(self, p):
         self.p = p
         self.n = self.w * self.p + self.b
         self.a = self.f(self.n)
         return
 
-    def learn(self, learning_rate, sensitivity):
-        self.w = self.w - learning_rate * sensitivity * np.transpose(self.p)
+    def update(self, learning_rate, sensitivity):
+        self.w = self.w - learning_rate * sensitivity * self.p.T
         self.b = self.b - learning_rate * sensitivity
         return
 
@@ -52,8 +52,7 @@ def tansig(n):
 
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
-    # a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-    return np.exp(n) / np.sum(np.exp(n), axis=0)
+    return np.exp(n) / np.sum(np.exp(n))
 
 # classify output
 def classify(a):
@@ -73,8 +72,8 @@ def j_softmax(a):
     for i in range(jacobian.shape[0]):
         for j in range(jacobian.shape[0]):
             if i == j:
-                x = np.zeros((a.shape[0]))
-                for k in range(jacobian.shape[0]):
+                x = np.zeros((a.shape[0], a.shape[0]))
+                for k in range(a.shape[0]):
                     x[k] = a[k] - a[i]
                 jacobian[i,j] = a[i] * np.sum(x)
             else:
@@ -87,25 +86,23 @@ def cross_entropy(a, t):
     i = np.where(a == 0)
     a[i] = 1e-15  # replace 0s so ln doesn't produce infinity
     e = -t * np.log(a)
-    #e = np.matrix([[-t[0,0] * np.log(a[0,0])], [-t[0,1] * np.log(a[1,0])]])
     return e
 
 
 # calculate hidden layer sensitivity
 def senseh(F_prime, W, s):
-    s = F_prime * np.transpose(W) * s
+    s = F_prime * W.T * s
     return s
 
 
 # calculate output layer sensitivity
-def senseo(F_prime, e):
-    s = -2 * F_prime * e
+def senseo(F_prime,t, a):
+    s = F_prime * (a - t)
     return s
 
-
-# define learning rule function
-def learn(weight_old, bias_old, sensitivity, input, learning_rate):
-    weight_new = weight_old - learning_rate * sensitivity * np.transpose(input)
+# update weights and biases
[…]

### 2017-06-29 Python Code/main_code.py

@@ -1,0 +1,138 @@
+# ----------------------------------------------------------------------------------------------------------------------
+# Import statements
+# ----------------------------------------------------------------------------------------------------------------------
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import NLObjects as MLP
+from NLObjects import cross_entropy
+from DataPrep import GlassImport
+from Split import split
+# ----------------------------------------------------------------------------------------------------------------------
+# Data pre-processing
+# ----------------------------------------------------------------------------------------------------------------------
+# import the glass dataset
+glass = GlassImport()
+# return the dimension of each input, R
+dim_inputs = glass.shape[1] - 2
+# split into training, validation, & testing sets
+train, validate, test = split(glass, dim_inputs, t=0.7, v=0.15)
+# number of neurons in layer 1
+num_neurons1 = 10
+# number of neurons in layer 2
+num_neurons2 = 2
+# learning rate
+alpha = 0.1
+# num_iter
+epoch = 150
+# instantiate layer 1
+nlayer1 = MLP.NeuronLayer(num_neurons1, dim_inputs, np.tanh, MLP.j_tansig)
+# instantiate layer 2
+nlayer2 = MLP.NeuronLayer(num_neurons2, num_neurons1, MLP.softmax, MLP.j_softmax)
+# set seed for reproducible results
+np.random.seed(150)
+# randomly initialize weight and bias on the interval [-0.5, 0.5]
+W1 = np.matrix(np.random.rand(num_neurons1, dim_inputs) - 0.5)
+b1 = np.matrix(np.random.rand(num_neurons1, 1) - 0.5)
+W2 = np.matrix(np.random.rand(num_neurons2, num_neurons1) - 0.5)
+b2 = np.matrix(np.random.rand(num_neurons2, 1) - 0.5)
+# pass the randomly initialized weight and bias onto the network object
+nlayer1.setWeightBias(w=W1, b=b1)
+nlayer2.setWeightBias(w=W2, b=b2)
+# ------------------- training set -------------------------
+# matrix of inputs
+p_train = train.iloc[:, range(dim_inputs)].T
+p_train = np.asmatrix(p_train)
+# matrix of targets
+target = train.iloc[:, -2:]
+target = np.asmatrix(target)
+# ------------------- validation set -------------------------
+# matrix of inputs
[…]

### 2017-06-29 Python Code/neuronLayer.py

@@ -1,13 +1,16 @@
+# ----------------------------------------------------------------------------------------------------------------------
+# import statements
+# ----------------------------------------------------------------------------------------------------------------------
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import NLObjects as nl
 import DataPrep as dp
 import Split as sp
-
+# ----------------------------------------------------------------------------------------------------------------------
+# Data pre-processing
+# ----------------------------------------------------------------------------------------------------------------------
 # normalize data and split dataset into training, validation, testing sets
-
-
 def main():
     # import dataset
     glass = dp.GlassImport()
@@ -19,7 +22,7 @@ def main():
     num_neurons1 = 10
     num_neurons2 = 2
     alpha = 0.1
-    epoch = 30
+    epoch = 300
     nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)
     nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)
     np.random.seed(0)
@@ -33,26 +36,19 @@ def main():
     # initialize cross entropy lists for training and validation sets
     ce_t = []
     ce_v = []
-
-    # training & validation
+    # Initialize weight and bias
     nlayer1.setWeightBias(w=W1, b=b1)
     nlayer2.setWeightBias(w=W2, b=b2)
-
+# ----------------------------------------------------------------------------------------------------------------------
+# training the network
+# ----------------------------------------------------------------------------------------------------------------------
     for j in range(epoch):
-
-
         for i in range(len(train)):
-
-            #set weights and biases
-
             # create input matrix from training dataset
             input = np.matrix(train.iloc[i, slice(0,num_inputs)]).transpose()
-
             # process network
-            nlayer1.processInput(input)
-            nlayer2.processInput(nlayer1.a)
-
-
+            nlayer1.FP(input)
+            nlayer2.FP(nlayer1.a)
             # calculate error of each iteration and update cost total
             target = np.matrix(train.iloc[i, -2:])
             e = nl.cross_entropy(nlayer2.a, target)
@@ -61,10 +57,9 @@ def main():
             else:
[…]

### 2017-06-29 Paper/paper.tex
Code edits

--- 2017-06-26 Paper/paper.tex
@@
 \label{sec4.2}
-%We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
+We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
 For multi-layer networks, the Levenberg-Marquardt algorithm is often used, but it does not work well for pattern recognition as the transfer function in the output layer is operating outside the linear region.

### 2017-06-29 Python Code/NLObjects.py

@@ -52,7 +52,7 @@ def tansig(n):
 
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
-    return np.exp(n) / np.sum(np.exp(n))
+    return np.exp(n) / np.sum(np.exp(n), axis=0)
 
 # classify output
 def classify(a):
@@ -96,8 +96,8 @@ def senseh(F_prime, W, s):
 
 
 # calculate output layer sensitivity
-def senseo(F_prime,t, a):
-    s = F_prime * (a - t)
+def senseo(t, a):
+    s = (a - t)
     return s
 
 # update weights and biases

### 2017-06-29 Python Code/neuronLayer.py

@@ -22,7 +22,7 @@ def main():
     num_neurons1 = 10
     num_neurons2 = 2
     alpha = 0.1
-    epoch = 300
+    epoch = 30
     nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)
     nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)
     np.random.seed(0)
@@ -59,7 +59,7 @@ def main():
             e = np.concatenate([e, e])
 
             # calculate layer 2 sensitivity
-            s2 = nl.senseo(F_prime=nlayer2.j(nlayer2.a), t=target, a=nlayer2.f(nlayer2.a))
+            s2 = nl.senseo(t=target, a=nlayer2.f(nlayer2.n))
             if i == 0:
                 s2_all = s2
             else:
@@ -119,7 +119,7 @@ def main():
     # plot error
     fig, ax = plt.subplots(ncols=1, nrows=1, figsize=[8, 8])
     ax.plot((np.arange(0, len(ce_t))), ce_t, label='Training')
-    ax.loglog((np.arange(0, len(ce_v))), ce_v, label='Validation')
+    ax.plot((np.arange(0, len(ce_v))), ce_v, label='Validation')
     ax.set_ylabel('Average Cross Entropy Error')
     ax.set_xlabel('Number of Batch Iterations')
     ax.legend()

### 2017-06-29 Python Code/main_code.py
Code Edits

@@ -5,7 +5,6 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import NLObjects as MLP
-from NLObjects import cross_entropy
 from DataPrep import GlassImport
 from Split import split
 # ----------------------------------------------------------------------------------------------------------------------
@@ -54,51 +53,54 @@ p_val = np.asmatrix(p_val)
 target_val = validate.iloc[:, -2:]
 target_val = np.asmatrix(target_val)
 # initialize cross entropy lists for training and validation sets
-e_train = []
-e_val = []
 ce_t = []
 ce_v = []
 # ----------------------------------------------------------------------------------------------------------------------
 # training the network
 # ----------------------------------------------------------------------------------------------------------------------
 for i in range(epoch):
-    s1_train = []
-    s2_train = []
     # ------------------- training set --------------------
     for q in range(len(train)):
-        # propagate the inputs forward
+        # propagate the inputs forward ---------------
         nlayer1.FP(p_train[:, q])
         nlayer2.FP(nlayer1.a)
-        # calculate the error
-        e_train.append(cross_entropy(nlayer2.a, target[q, :]))
-        # backprop
-        s2_train.append(MLP.senseo(nlayer2.j(nlayer2.a), target[q, :].T, nlayer2.f(nlayer2.a)))
-        s1_train.append(MLP.senseh(nlayer1.j(nlayer1.a), nlayer2.w, s2_train[:, q]))
-    # update weight and bias
-    s1_train_epoch = np.mean(s1_train)
-    s2_train_epoch = np.mean(s2_train)
-    nlayer2.update(sensitivity=s2_train_epoch, learning_rate=alpha)  # layer 1
-    nlayer1.update(sensitivity=s1_train_epoch, learning_rate=alpha)  # layer 2
-    ce_t.append(np.mean(e_train))
+        # calculate the error ------------
+        ce = MLP.cross_entropy(nlayer2.a, target[q, :])
+        if i == 0:
+            e_all = ce
+        else:
+            e_all = np.concatenate((e_all, ce), axis=1)
+        ce = np.concatenate([ce, ce])
+        # backprop ----------------
+        # layer 2 sensitivity
[…]

### 2017-06-29 Python Code/neuronLayer.py

@@ -10,7 +10,8 @@ import Split as sp
 # ----------------------------------------------------------------------------------------------------------------------
 # Data pre-processing
 # ----------------------------------------------------------------------------------------------------------------------
-# normalize data and split dataset into training, validation, testing sets
+
+
 def main():
     # import dataset
     glass = dp.GlassImport()
@@ -36,6 +37,7 @@ def main():
     # initialize cross entropy lists for training and validation sets
     ce_t = []
     ce_v = []
+    global s1_all, s2_all
     # Initialize weight and bias
     nlayer1.setWeightBias(w=W1, b=b1)
     nlayer2.setWeightBias(w=W2, b=b2)
@@ -45,7 +47,7 @@ def main():
     for j in range(epoch):
         for i in range(len(train)):
             # create input matrix from training dataset
-            input = np.matrix(train.iloc[i, slice(0,num_inputs)]).transpose()
+            input = np.matrix(train.iloc[i, slice(0, num_inputs)]).transpose()
             # process network
             nlayer1.FP(input)
             nlayer2.FP(nlayer1.a)
@@ -81,7 +83,7 @@ def main():
         nlayer1.update(sensitivity=s1_all.mean(axis=1), learning_rate=alpha)
 
         # validation
-        input = np.matrix(validate.iloc[:, slice(0,num_inputs)]).transpose()
+        input = np.matrix(validate.iloc[:, slice(0, num_inputs)]).transpose()
         # first layer
         nlayer1.FP(input)
         nlayer2.FP(nlayer1.a)
@@ -93,35 +95,52 @@ def main():
                 e_all = e
             else:
                 e_all = np.concatenate((e_all, e), axis=1)
-
-        # append average validation cross entropy to list
         ce_v.append(e_all.mean())
-        if j >= 5 and ce_v[j] > ce_v[j - 1]:
-            break
+        # early stopping
+        if j == 0:
+            val_fail = []
+        elif ce_v[j] > ce_v[j-1]:
+            val_fail.append(1)
+            if len(val_fail) == 5:
+                print 'Validation error has increased for 5 consecutive epochs. Early stopping at epoch {}'.format(j)
+                break
+        else:
+            val_fail = []
 # ----------------------------------------------------------------------------------------------------------------------
[…]

### 2017-06-30 Paper/paper.tex
Code edits

@@
 
-%Widrow and Hoff's learning rule suffered from the same disadvantage.
-When we have more complicated decision boundaries we are trying to implement and have multiple layers, a more robust learning rule ought to be used.
-Since then, multilayer perceptrons trained by the backpropagation algorithm have become the most popular neural network.
-\cite{hagan2014}.
+Widrow and Hoff's Least Mean Square (LMS) learning rule suffered from the same disadvantage as Rosenblatt's, but it has since been generalized.
+The generalization of the LMS learning rule is referred to as \emph{backpropagation} and we commonly use it to train multi-layer perceptron networks \cite{hagan2014}.
 
@@
 \label{sec4.2}
-We chose to use the Scaled Conjugate Gradient (SCG) algorithm to train our network as it is very efficient for pattern recognition problems.
+We chose to use the Scaled Conjugate Gradient (SCG) algorithm in MATLAB to train our network as it is very efficient for pattern recognition problems.
 For multi-layer networks, the Levenberg-Marquardt algorithm is often used, but it does not work well for pattern recognition as the transfer function in the output layer is operating outside the linear region.
 The scaled conjugate gradient algorithm is a special type of backpropagation.
-In backpropagation, we adjust the weights in the direction of steepest descent...
+For the replication of the network in Python, we use standard backpropagation.
+
+The implementation of backpropagation can be broken down into three steps: 
+\begin{enumerate}
+\item Propagate the input forward through the network
+\item Propagate the sensitivities backward through the network
+\item Update the weights and biases using the approximate steepest descent rule
+\end{enumerate}
+
+\subsubsection{Forward Propagation}
 

### 2017-06-30 Python Code/neuronLayer.py

@@ -13,32 +13,29 @@ import Split as sp
 
 
 def main():
-    # import dataset
+    # --------------------- import data ---------------------
     glass = dp.GlassImport()
     num_inputs = glass.shape[1] - 2
-    # split into training, validate, testing sets
+    # split into training, validation & testing data
     train, validate, test = sp.split(glass, num_inputs, t=0.7, v=0.15)
-
-    # specify number of neurons in each layer and the learning rate
-    num_neurons1 = 10
-    num_neurons2 = 2
-    alpha = 0.1
-    epoch = 30
-    nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)
-    nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)
+    # --------------------- specify network architecture ---------------------
+    num_neurons1 = 10  # layer 1
+    num_neurons2 = 2   # layer 2
+    alpha = 0.01       # learning rate
+    epoch = 100        # number of iterations
+    nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)      # instantiate layer 1
+    nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)  # instantiate layer 2
     np.random.seed(0)
-    # initialize weight and bias randomly for each layer from -0.5 to 0.5
-    # W1 number of columns matches training set columns, less final two
+    # randomly initialize weight and bias on the interval [-0.5, 0.5]
     W1 = np.matrix(np.random.rand(num_neurons1, num_inputs) - 0.5)
     b1 = np.matrix(np.random.rand(num_neurons1, 1) - 0.5)
     W2 = np.matrix(np.random.rand(num_neurons2, num_neurons1) - 0.5)
     b2 = np.matrix(np.random.rand(num_neurons2, 1) - 0.5)
-
-    # initialize cross entropy lists for training and validation sets
+    # initialize cross entropy loss (training & validation)
     ce_t = []
     ce_v = []
     global s1_all, s2_all
-    # Initialize weight and bias
+    # pass on randomly initialized weights and biases to the network
     nlayer1.setWeightBias(w=W1, b=b1)
     nlayer2.setWeightBias(w=W2, b=b2)
 # ----------------------------------------------------------------------------------------------------------------------
@@ -48,47 +45,41 @@ def main():
         for i in range(len(train)):
             # create input matrix from training dataset
[…]

### 2017-06-30 Paper/paper.tex
Paper Edits

@@
 They offer up the following example of a rule:
-\[ 
+\begin{equation}
 \{(\mathrm{Fe} \leq \mathrm{Na}) \text{ and } [\mathrm{K} > (\mathrm{Fe} \cdot 650)]\}, 
-\]
+\end{equation}
 where $\mathrm{Fe}$, $\mathrm{Na}$, and $\mathrm{K}$ are the percent composition of iron, sodium, and potassium, respectively, of each glass sample.
@@
 Now that we have chosen a network architecture, we can calculate the network output. The output from the hidden layer (the input to the output layer) can be calculated as 
-\[
+\begin{equation}
 \mathbf{a^{1}} = \textbf{tansig}(\mathbf{W^{1}}\mathbf{p} + \mathbf{b^{1}}), 
-\] 
+\end{equation}
 while the output from the output layer is 
-\[ 
+\begin{equation}
 \mathbf{a^{2}} = \textbf{softmax}(\mathbf{W^{2}}\mathbf{a^{1}} + \mathbf{b^{2}}). 
-\]
+\end{equation}
 The network architecture can be seen in \textbf{Figure 1} below. 
@@
 The scaled conjugate gradient algorithm is a special type of backpropagation.
-For the replication of the network in Python, we use standard backpropagation.
+For the replication of the network in Python, we use standard backpropagation with steepest descent.
 
@@
 \subsubsection{Forward Propagation}
+\label{sec4.2.1}
+For a multilayer network, the output from one layer is the input to the next layer.
+That is, the hidden layer's output is the input to the output layer of the network.
+The neurons in the first layer receive external inputs (the observed data):
+\begin{equation}
+\mathbf{a}^{0} = \mathbf{p},
+\end{equation}
+which serves as the starting point for the network. After the input is received, it is propagated forward with the following equation
+\begin{equation}
+\mathbf{a}^{m+1} = \mathbf{f}^{m+1}(\mathbf{W}^{m+1}\mathbf{a}^{m} + \mathbf{b}^{m+1}), \text{ for \emph{m} = 0, 1, \ldots, $M - 1$} 
+\end{equation}
+where we use $M$ to denote the number of layers in the network. Then, the output from the the neurons in the last layer are
+\begin{equation}
+\mathbf{a} = \mathbf{a}^{M}.
+\end{equation}
+
+\subsubsection{Performance Index}
+\label{sec4.2.2}
+Before moving on to the step of backpropagating the sensitivities, we discuss the performance index as it is a critical part of the process.
+The performance index is a measure of the error of the network outputs in relation to the targets.
+The general implementation of the backpropagation algorithm uses the \emph{mean square error} as a performance index.
+The mean square error is approximated by taking the expectation of the sum of the squared errors (residuals).
+However, we choose a different performance index to gauge our model's predictions.
+
+Mean square error works very well for functions with continuous target values---the case where we are approximating a function.
+However, in pattern recognition we are given discrete target values, so other performance indices that take this into account make more sense.
+We choose to use \emph{cross-entropy} as our performance index.
+This is a commonly used performance index when the \emph{softmax} transfer function is used in the output layer.
+Given a set of input-target pairs
+\begin{equation}
+\{\mathbf{p}_{1}, \mathbf{t}_{1}\}, \{\mathbf{p}_{2}, \mathbf{t}_{2}\}, \ldots, \{\mathbf{p}_{Q}, \mathbf{t}_{Q}\}
+\end{equation}
+where $\mathbf{p}_{q}$ is an input and $\mathbf{t}_{q}$ is the corresponding target output, we denote the cross-entropy loss as follows: 
+\begin{equation}
+F(\mathbf{x}) = - \sum_{q = 1}^{Q} \sum_{i = 1}^{S^{M}} t_{i, q} \ln \frac{a_{i, q}}{t_{i, q}}
+\end{equation}
+where $Q$ is the number of samples in the dataset and $S^{M}$ is the number of neurons in the output layer.
+We can simplify this by vectorizing the operation over all input-target pairs, $\{\mathbf{p}_{q}, \mathbf{t}_{q}\}$, and eliminating the first summation:
+\begin{equation}
+- \sum_{i = 1}^{S^{M}} \mathbf{t}_{i} \ln \frac{\mathbf{a}_{i}}{\mathbf{t}_{i}}.
+\end{equation}
+Recall that in our pattern recognition problem we have two classes where the targets are $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$ and $\begin{bmatrix} 0 \\ 1 \end{bmatrix}$.
+Thus, each neuron can only take on values of zero or one.
+This allows us to further simplify the equation to
+\begin{equation}
+- \mathbf{t} \ln \mathbf{a}
+\end{equation}
+since $\mathbf{t}_{i} = 0$ implies the cross-entropy loss for the $i$-th neuron is zero and the case where 
+$\mathbf{t}_{i} = 1$ implies the cross-entropy loss is simply $ - \mathbf{t}_{i} \ln \mathbf{a}_{i}$.
+
+It is important to note the backpropagation algorithm works with any differentiable performance index we specify.
+We just need to change the initialization of the sensitivities in the output layer accordingly.
+This brings us back to the next step: propagating the sensitivities backward through the network \cite{hagan2014}.
+
+\subsubsection{Backpropagation}
+\label{sec4.2.3}
+Once we have propagated the input forward, we tcompare the network output to the targets so we can use the error to adjust the weights and biases accordingly.
+But, in the case of a multi-layer network, the errors and the performance index we use to evaluate those errors are no longer just a function of the weights.
+Rather, they are indirectly a function of the weights in the hidden layer in addition to the weights in the output layer.
+Thus, we call the errors we propagate backward through the network `sensitivities' because they represent the sensitivity of the performance index, $F(\mathbf{x})$, to changes in the $i$-th element of the net input at layer $m$.
+We can calculate these sensitivities analytically as follows:
+\begin{equation}
+s^{m}_{i} \equiv \frac{\partial F}{\partial n^{m}_{i}}. 
+\end{equation}
+Now, we compute the sensitivities $\mathbf{s}^{m}$.
+Note that the reason we call it backpropagation is because the sensitivities are propagated backward through the network via a recurrence relation where the sensitivity at layer $m$ is calculated from the sensitivity at layer $m+1$.
+This is easier said than done, however, since we need to find a Jacobian matrix of sensitivities to derive the recurrence relation:
+\begin{equation}
+\frac{\partial \mathbf{n}^{m+1}}{\partial \mathbf{n}^{m}} \equiv \begin{bmatrix} 
+\frac{\partial n^{m+1}_{1}}{\partial n^{m}_{1}} &  \frac{\partial n^{m+1}_{1}}{\partial n^{m}_{2}} & \dots &  \frac{\partial n^{m+1}_{1}}{\partial n^{m}_{S^{m}}} \\
+\frac{\partial n^{m+1}_{2}}{\partial n^{m}_{1}} &  \frac{\partial n^{m+1}_{2}}{\partial n^{m}_{2}} & \dots & \frac{\partial n^{m+1}_{2}}{\partial n^{m}_{S^{m}}} \\
+\vdots & \vdots &  & \vdots \\
+\frac{\partial n^{m+1}_{S^{m+1}}}{\partial n^{m}_{1}} &  \frac{\partial n^{m+1}_{S^{m+1}}}{\partial n^{m}_{2}} & \dots & \frac{\partial n^{m+1}_{S^{m+1}}}{\partial n^{m}_{S^{m}}}
+\end{bmatrix}
+\end{equation}
+To find an expression to represent this Jacobian matrix, we arbitrarily select the $i, j$ element of the matrix in accordance with Hagan's derivation: 
+\begin{equation}
+\begin{split}
+\frac{\partial n^{m+1}_{i}}{\partial \mathbf{n}^{m}_{j}} &= \frac{\partial \bigg(\sum_{l=1}^{S^{m}} w_{i, l}^{m+1}a_{l}^{m} + b_{i}^{m+1}\bigg)}{\partial n_{j}^{m}} = w_{i, j}^{m+1} \frac{\partial a_{j}^{m}}{\partial n_{j}^{m}} \\
+&= w_{i, j}^{m+1} \frac{\partial f^{m}(n_{j}^{m}}{\partial n_{j}^{m}} = w_{i, j}^{m+1} \dot{f}^{m}(n_{j}^{m}) ,  
+\end{split}
+\end{equation}
+where 
+\begin{equation}
+\dot{f}^{m}(n_{j}^{m}) = \frac{\partial f^{m}(n_{j}^{m})}{\partial n_{j}^{m}}.
+\end{equation}
+Thus, we can write the Jacobian from before as 
+\begin{equation}
+\frac{\partial \mathbf{n}^{m+1}}{\partial \mathbf{n}^{m}} = \mathbf{W}^{m+1} \dot{\mathbf{F}}^{m}(\mathbf{n}^{m}), 
+\end{equation}
+where 
+\begin{equation}
+\dot{\mathbf{F}}^{m}(\mathbf{n}^{m}) = \begin{bmatrix} \dot{f}^{m}(n_{1}^{m}) & 0 & \ldots & 0 \\ 
+0 & \dot{f}^{m}(n_{2}^{m}) & \ldots & 0 \\
+\vdots & \vdots & \ddots & \vdots \\ 
+0 & 0 & \ldots & \dot{f}^{m}(n_{S^{m}}^{m}) \end{bmatrix}.
+\end{equation}
+Now the recurrence relation for the hidden layer sensitivities can be written as 
+\begin{equation}
+\begin{split}
+\mathbf{s}^{m} & = \dot{\mathbf{F}}^{m}(\mathbf{n}^{m})(\mathbf{W}^{m+1})^{T}\frac{\partial \hat{F}}{\partial \mathbf{n}^{m+1}} \\
+& = \dot{\mathbf{F}}^{m}(\mathbf{n}^{m})(\mathbf{W}^{m+1})^{T}\mathbf{s}^{m+1},
+\end{split}
+\end{equation}
+where $\hat{F}$ denotes the performance index as it is an approximation of the network error \cite{hagan2014}.
+But, we know the network architecture and performance index used, so we can refine the general equation.
+The transfer function in the hidden layer is the tangent-sigmoid function and we find the derivative as follows:
+\begin{equation}
+\begin{split}
+\dot{f}^{1} & = \frac{d(\tanh(n))}{dn} = \frac{d}{dn}\bigg(\frac{\sinh(n)}{\cosh(n)}\bigg) \\
+ & = \frac{\cosh^{2}(n) - \sinh^{2}(n)}{\cosh^{2}(n)} = 1 - \tanh^{2}(n). 
+\end{split}
+\end{equation}
+Therefore, 
+\begin{equation}
+\dot{\mathbf{F}}^{1}(\mathbf{n}^{1}) = \begin{bmatrix} 1 - \tanh^{2}(n_{1}^{1}) & 0 & \ldots & 0 \\ 
+0 & 1 - \tanh^{2}(n_{2}^{1}) & \ldots & 0 \\
+\vdots & \vdots & \ddots & \vdots \\ 
+0 & 0 & \ldots & 1 - \tanh^{2}(n_{10}^{1}) \end{bmatrix}.
+\end{equation}
+and the first layer sensitivity is 
+\begin{equation}
+\mathbf{s}^{1} = \dot{\mathbf{F}}^{1}(\mathbf{n}^{1})(\mathbf{W}^{2})^{T}\mathbf{s}^{2}.
+\end{equation}
+The only thing left to do is calculate the sensitivity for the output layer, $\mathbf{s}^{2}$.
+Calculating the sensitivity for the output layer is a little more difficult as the derivative of the performance index with respect to the net input of the output layer is indirectly a function of the net input of the hidden layer.
+We can derive the sensitivity for the output layer as follows:
+\begin{equation}
+\begin{split}
+\mathbf{s}^{2} & = \frac{\partial \hat{F}}{\partial \mathbf{n}_{i}^{2}} = - \sum_{s = 1}^{2} \frac{\partial \mathbf{t}_{s} \ln \mathbf{a}_{s}}{\partial \mathbf{n}_{s}} \\
+& = - \sum_{s = 1}^{2} \mathbf{t}_{s} \frac{\partial \ln \mathbf{a}_{s}}{\partial \mathbf{n}_{i}} = - \sum_{s = 1}^{2} \mathbf{t}_{s} \frac{1}{\mathbf{a}_{s}} \frac{\partial \mathbf{a}_{s}}{\partial \mathbf{n}_{i}}
+\end{split}
+\end{equation}
+where $i$ and $s$ denote the $i$-th and $s$-th neuron, respectively, and 
+\begin{equation}
+\frac{\partial a_{s}}{\partial n_{i}} = \dot{\mathbf{F}}^{2}.
+\end{equation}
+Now we find the derivative of the \emph{softmax} transfer function to be used in the sensitivity calculation: 
+\begin{equation}
+\begin{split}
+\text{if } i = j\text{: } \frac{\partial a_{i}}{\partial n_{i}} = \frac{\partial \textbf{softmax}(n_{i})}{\partial n_{i}}
+
+\text{if } i \neq j\text{: }
+\end{split}
+\end{equation}
 

### 2017-06-30 Python Code/NLObjects.py
Code test edits

@@ -72,10 +72,11 @@ def j_softmax(a):
     for i in range(jacobian.shape[0]):
         for j in range(jacobian.shape[0]):
             if i == j:
-                x = np.zeros((a.shape[0], a.shape[0]))
-                for k in range(a.shape[0]):
-                    x[k] = a[k] - a[i]
-                jacobian[i,j] = a[i] * np.sum(x)
+                #x = np.zeros((a.shape[0], a.shape[0]))
+                #for k in range(a.shape[0]):
+                    #x[k] = a[k] - a[i]
+                #jacobian[i,j] = a[i] * np.sum(x)
+                jacobian[i,j] = a[i] * (1 - a[i])
             else:
                 jacobian[i, j] = -a[i] * a[j]
     return jacobian

### 2017-06-30 Python Code/NLObjects.py
Fix Gradient of Softmax

@@ -72,10 +72,6 @@ def j_softmax(a):
     for i in range(jacobian.shape[0]):
         for j in range(jacobian.shape[0]):
             if i == j:
-                #x = np.zeros((a.shape[0], a.shape[0]))
-                #for k in range(a.shape[0]):
-                    #x[k] = a[k] - a[i]
-                #jacobian[i,j] = a[i] * np.sum(x)
                 jacobian[i,j] = a[i] * (1 - a[i])
             else:
                 jacobian[i, j] = -a[i] * a[j]

### 2017-06-30 Python Code/neuronLayer.py

@@ -22,7 +22,7 @@ def main():
     num_neurons1 = 10  # layer 1
     num_neurons2 = 2   # layer 2
     alpha = 0.01       # learning rate
-    epoch = 50        # number of iterations
+    epoch = 100        # number of iterations
     nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)      # instantiate layer 1
     nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)  # instantiate layer 2
     np.random.seed(0)
@@ -109,7 +109,7 @@ def main():
     predict = np.array(predict).flatten()
     predict = pd.Series(predict, name='Predicted')
     confusion = pd.crosstab(actual, predict, margins=False)  # create confusion matrix
-    confusion = confusion.astype(float) # convert values to floats
+    confusion = confusion.astype(float)  # convert values to floats
     print confusion  # output confusion matrix to the console
     # --------------------- accuracy metrics ---------------------
     ERR = (confusion.loc[0, 1] + confusion.loc[1, 0]) / len(predict)

### 2017-06-30 Python Code/neuronLayer.py
Add sensitivity derivation

@@ -135,7 +135,7 @@ def main():
     fig, ax = plt.subplots(ncols=1, nrows=1, figsize=[8, 8])
     ax.plot((np.arange(0, len(ce_t))), np.log(ce_t), label='Training', linewidth=2.0, color='blue')
     ax.plot((np.arange(0, len(ce_v))), np.log(ce_v), label='Validation', linewidth=2.0, color='green')
-    ax.set_ylabel('Log Cross Entropy Error')
+    ax.set_ylabel('Log Cross Entropy Loss')
     ax.set_xlabel('Epochs')
     ax.legend()
     plt.grid()

### 2017-06-30 Paper/paper.tex
Add sensitivity derivation
Add Appendices

@@
 \documentclass[12pt,halfline,a4paper]{ouparticle}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{amsfonts}
+\usepackage{listings}
+\usepackage{tikz}
+\usepackage{fancyhdr}
+\usepackage{extramarks}
+\usepackage{algpseudocode}
+\usetikzlibrary{automata,positioning}
 \usepackage[utf8]{inputenc}
@@
 \usepackage[nottoc]{tocbibind}
-
 \addbibresource{paper.bib}
 
+% defining colors chunks 
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+% define a global listing style
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=3pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=1
+}
+
+% set global listing style
+\lstset{style=mystyle}
+
+% define styles for each language
+
+\lstdefinestyle{R}
+    {language = R, style = mystyle}
+\lstdefinestyle{Python}
+    {language = Python, style = mystyle}
+\lstdefinestyle{Matlab}
+    {language = Matlab, style = mystyle}
+
 \begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% title page
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% table of contents
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \pagebreak
@@
 \pagebreak
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% paper
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{split}
-\mathbf{s}^{2} & = \frac{\partial \hat{F}}{\partial \mathbf{n}_{i}^{2}} = - \sum_{s = 1}^{2} \frac{\partial \mathbf{t}_{s} \ln \mathbf{a}_{s}}{\partial \mathbf{n}_{s}} \\
-& = - \sum_{s = 1}^{2} \mathbf{t}_{s} \frac{\partial \ln \mathbf{a}_{s}}{\partial \mathbf{n}_{i}} = - \sum_{s = 1}^{2} \mathbf{t}_{s} \frac{1}{\mathbf{a}_{s}} \frac{\partial \mathbf{a}_{s}}{\partial \mathbf{n}_{i}}
+\mathbf{s}^{2} & = \frac{\partial \hat{F}}{\partial \mathbf{n}_{i}^{2}} = - \sum_{i = 1}^{2} \frac{\partial \mathbf{t}_{j} \ln \mathbf{a}_{j}}{\partial \mathbf{n}_{i}} \\
+& = - \sum_{j = 1}^{2} \mathbf{t}_{j} \frac{\partial \ln \mathbf{a}_{j}}{\partial \mathbf{n}_{i}} = - \sum_{j = 1}^{2} \mathbf{t}_{j} \frac{1}{\mathbf{a}_{j}} \frac{\partial \mathbf{a}_{j}}{\partial \mathbf{n}_{i}}
 \end{split}
 \end{equation}
-where $i$ and $s$ denote the $i$-th and $s$-th neuron, respectively, and 
+where $i$ denotes the $i$-th neuron and 
 \begin{equation}
-\frac{\partial a_{s}}{\partial n_{i}} = \dot{\mathbf{F}}^{2}.
+\frac{\partial a_{i}}{\partial n_{i}} = \dot{\mathbf{F}}^{2}.
 \end{equation}
@@
 \begin{split}
-\text{if } i = j\text{: } \frac{\partial a_{i}}{\partial n_{i}} = \frac{\partial \textbf{softmax}(n_{i})}{\partial n_{i}}
+\text{if } i = j \text{: } \frac{\partial a_{i}}{\partial n_{i}} & = \frac{\partial \textbf{softmax}(n_{i})}{\partial n_{i}} = \frac{e^{n_{i}}}{\sum_{n = 1}^{2} e^{n_{i}}} \bigg(\frac{\sum_{n = 1}^{2} e^{n_{i}} - e^{n_{i}}}{\sum_{n = 1}^{2} e^{n_{i}}}\bigg) \\
+& = \frac{e^{n_{i}}}{\sum_{n = 1}^{2} e^{n_{i}}}\bigg(1 - \frac{e^{n_{i}}}{\sum_{n = 1}^{2} e^{n_{i}}}\bigg) = a_{i}(1 - a_{i}) \\ \\
+\text{if } i \neq j \text{: } \frac{\partial a_{i}}{\partial n_{j}} & = \frac{\partial \textbf{softmax}(n_{i})}{\partial n_{j}} = - \frac{e^{n_{i}}}{\sum_{n = 1}^{2} e^{n_{i}}} \frac{e^{n_{j}}}{\sum_{n = 1}^{2} e^{n_{j}}} = -a_{i}a_{j}.
+\end{split}
+\end{equation}
+Using this result to finish the derivation of $\mathbf{s}^{2}$ from (21), we have: 
+\begin{equation}
+\begin{split}
+s_{i}^{2} &= - \sum_{j = 1}^{2} \mathbf{t}_{j} \frac{1}{\mathbf{a}_{j}} \frac{\partial \mathbf{a}_{j}}{\partial \mathbf{n}_{i}} = - \frac{\mathbf{t}_{i}}{\mathbf{a}_{i}} \frac{\partial \mathbf{a}_{i}}{\partial \mathbf{n}_{i}} - \sum_{j \neq i}^{2} \frac{\mathbf{t}_{j}}{\mathbf{a}_{j}} \frac{\partial \mathbf{a}_{j}}{\partial \mathbf{n}_{i}} \\
+& = - \frac{\mathbf{t}_{i}}{\mathbf{a}_{i}} \mathbf{a}_{i}(1 - \mathbf{a}_{i}) - \sum_{j \neq i}^{2} \frac{\mathbf{t}_{j}}{\mathbf{a}_{j}}(-a_{i}a_{j}) \\
+& = - \mathbf{t}_{i} + \mathbf{t}_{i}\mathbf{a}_{i} + \sum_{j \neq i}^{2} \mathbf{t}_{j}\mathbf{a}_{i} = - \mathbf{t}_{i} + \sum_{j = 1}^{2} \mathbf{t}_{j}\mathbf{a}_{i} \\
+& = - \mathbf{t}_{i} + \mathbf{a}_{i} \sum_{j = 1}^{2} \mathbf{t}_{j} = \mathbf{a}_{i}  - \mathbf{t}_{i}.
+\end{split}
+\end{equation}
+Now we have the starting sensitivity: 
+\begin{equation}
+\mathbf{s}^{2} = \mathbf{a} - \mathbf{t}
+\end{equation}
+and have everything we need to backpropagate the sensitivities for our network.
 
-\text{if } i \neq j\text{: }
-\end{split}
+\subsubsection{Update Rule}
+\label{sec4.2.4}
+Lastly, we need to update the weights and biases via an update rule. For the replication in Python, we use the approximate steepest descent rule to update weights and biases: 
+\begin{equation}
+\mathbf{W}^{m}(k+1) = \mathbf{W}^{m}(k) - \alpha \mathbf{s}^{m}(\mathbf{a}^{m-1})^{T}, 
+\end{equation}
+\begin{equation}
+\mathbf{b}^{m}(k+1) = \mathbf{b}^{m}(k) - \alpha \mathbf{s}^{m}. 
 \end{equation}
@@
 \label{sec5}
+This section will document how the data are used to train and test the network, implementation of the model in Python, and performance metrics used to evaluate the final model.
 
@@
 \label{sec5.1}
+The data were almost ready to be used when we downloaded it as there were no missing values.
+The only thing that needed to be changed from the original dataset was the target values.
+Originally, these values were coded on a one-to-seven scale where values of one-to-four denoted a type of window glass and values of five-to-seven denoted non-window glass.
+When coding the targets for pattern recognition problems, it is best practice to have as many dimensions in the targets as there are classes.
+This ensures only one neuron is active at a time and generally produces better results.
+Further, we coded the targets as $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$ if the glass sample comes from a window and $\begin{bmatrix} 0 \\ 1 \end{bmatrix}$ it it does not for this reason.
+We briefly mention this in Section \ref{sec3.2} during our discussion of the targets.
+The code used to import the data and code the targets can be found in Appendix A.
+
+After collecting the data, we normalize it based on the network architecture we have chosen.
+Since we use a tangent-sigmoid transfer function in the hidden layer, it is very important that we normalize the inputs between negative one and one since this is the range of the transfer function.
+When the net input is large (or small), the function becomes saturated.
+We do not want this to happen at the beginning of the training process because the gradient will be close to zero.
+Since the first layer's net input is the product of the weight and input plus the bias, large inputs can result in large net inputs if the weights are not small enough scale the inputs down.
+Thus, it is important to normalize the inputs before we feed them to the network.
+In order to normalize the inputs to values between negative one and one, we use the following equation:
+\begin{equation}
+\mathbf{p}^{n} = 2(\mathbf{p}-\mathbf{p}^{min}) ./ (\mathbf{p}^{max}-\mathbf{p}^{min}) - 1,
+\end{equation}
+where $\mathbf{p}^{min}$ is the vector containing the minimum values of each element of the input vectors in the dataset, $\mathbf{p}^{max}$ contains the maximum values, ./ represents the element-wise division of the vectors, and $\mathbf{p}^{n}$ is the normalized input vector \cite{hagan2014}.
+
+Next, we divide the data up into three different sets---training, validation, and testing.
+We chose to use 70\% of the data for training, 15\% of the data for validation, and 15\% of the data for testing.
+This is fairly standard practice and follows the literature.
+To ensure that each of these sets is representative of the full dataset, we split the the data up into two subsets---input-target pairs for window samples and input-target pairs for non-windows.
+
+Then, we split each of these subsets up into training, validation, and testing sets using the 70:15:15 ratio.
+After splitting the subsets up into training, validation, and testing, we combine each split-up subset accordingly---i.e.
+window-training subset with non-window-training subset.
+This gives us the highest chance of having our subsets be representative of the full dataset and that the subsets cover the same region of the input space.
+Once the subsets have been combined into complete training, validation, and testing sets, we reshuffle the input-target pairs so that they are randomly-ordered again.
+The subroutine used to normalize the inputs and split the data can be found in Appendix B.
 
@@
 \label{sec5.2}
+To implement the network, we created a Python class, \verb|NeuronLayer|, to represent layers of a neural network.
+The \verb|NeuronLayer| class takes the dimension of the inputs, number of neurons in the layer, transfer function, and gradient of the transfer function as arguments.
+Moreover, there are several functions defined to help train and implement the network in Python.
+The subroutine used to create the \verb|NeuronLayer| class can be found in Appendix C.
 
-\subsection{Performance Index}
-\label{sec5.3}
+We begin by instantiating layers from the created class to mimic the network architecture we designed.
+We arbitrarily chose a learning rate, $\alpha$, of 0.01.
+If the learning rate converges too slowly, then we will increase the learning rate accordingly.
+Then, we randomly initialize the weights and biases on the interval $[-0.5, 0.5]$ and pass these on to the corresponding network layers with the \verb|setWeightBias| method.
+Weights and biases are usually initialized on this interval when we have normalized our inputs between negative one and one.
+Lastly, we initialize empty vectors to hold the error from each training iteration.
+These steps can be found in Appendix D.
+
+\subsection{Training the Network}
+We chose to train the network over 100 epochs where one epoch represents a single batch implementation of approximate steepest descent.
+The training process begins by propagating each input in the training set forward by calling the \verb|FP| method of the \verb|NeuronLayer| class.
+Then, the cross-entropy is calculated using the equation found in (8).
+This error is then used to calculate a sensitivity and then propagated backward through the model.
+Once this process is done for each input in the dataset, we approximate the cross-entropy loss of the epoch by taking the mean of the cross-entropy of the given epoch.
+Then, we use this information in addition to the learning rate to update the weights and biases by (26) and (27), respectively.
+
+In attempts to speed up the iterative process used to train the model, we perform the validation simultaneously.
+We use the same process to propagate inputs forward and compute the cross-entropy loss, but instead of updating the weights and biases based on the validation set, we use the weights and biases from the training set.
+This allows us to see if the network is training properly since the validation set is information our model has not been trained on.
+If the error keeps decreasing in the validation set as the number of iterations increases, this means our network is training properly.
+To ensure we do not overfit the data with our model, we implement an early stopping technique.
+We decided to have our network stop training if the validation error increased for five \emph{consecutive} epochs.
+If this happens to be the case, the script for the model will print to the console that training was stopped early and at what iteration it stopped.
+The code used to implement this can be found in Appendix E or in the include file \verb|NeuronLayer.py|.
 
@@
 
-\section{Perceptron}
-\section{Backprop}
-\section{Subroutine X}
+\section{Importing the Data \& Coding the Targets}
+
+The Python code used to import the data and code the targets is listed below.
+First the data are imported with a standard \verb|read_csv| statement from the \verb|pandas| library.
+Then, the targets are coded by the process stated in Section 3.2.
+The full subroutine can be found in the included file \verb|DataPrep.py|.
+
+\begin{lstlisting}[style = Python]
+def GlassImport():
+    # import dataset
+    cols = ['id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'type']
+    glass = pd.read_csv('glass.txt', names=cols)
+    # create window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] <= 4, 'window'] = 1
+    glass.ix[glass.ix[:, 'type'] > 4, 'window'] = 0
+    # create non-window target column (1 for true; 0 for false)
+    glass.ix[glass.ix[:, 'type'] > 4, 'non_window'] = 1
+    glass.ix[glass.ix[:, 'type'] <= 4, 'non_window'] = 0
+    # drop id and class columns
+    glass.drop(['id', 'type'], inplace=True, axis=1)
+    return glass
+\end{lstlisting} 
+\pagebreak
+
+\section{Data Normalization \& Division}
+The following code documents the process used to normalize the inputs and then divide the data up into training, validation, and testing sets.
+The full subroutine can be found in the included file \verb|Split.py|.
+
+\begin{lstlisting}[style = Python]
+def split(dataset, num_inputs, t, v):
+    # normalize dataset into range between -1 and 1
+    norm = dataset.ix[:, slice(0,num_inputs)].apply(lambda x: -1 + 2 * (x - x.min()) / (x.max() - x.min()), axis=0)
+    dataset = pd.concat([norm, dataset.ix[:,'window':]], axis=1)
+    # split dataset by target class
+    dataset0 = dataset.ix[dataset.ix[:, 'window'] == 0, :]
+    dataset1 = dataset.ix[dataset.ix[:, 'window'] == 1, :]
+    # build datasets by target class
+    train0, validate0, test0 = np.split(dataset0.sample(frac=1),
+                                        [int(t * len(dataset0)), int((1 - v) * len(dataset0))])
+    train1, validate1, test1 = np.split(dataset1.sample(frac=1),
+                                        [int(t * len(dataset1)), int((1 - v) * len(dataset1))])
+    # re-combine by target class
+    train = pd.concat([train0, train1])
+    validate = pd.concat([validate0, validate1])
+    test = pd.concat([test0, test1])
+    # shuffle order and reset index numbers
+    train = train.sample(frac=1)
+    train.reset_index(inplace=True)
+    validate = validate.sample(frac=1)
+    validate.reset_index(inplace=True)
+    test = test.sample(frac=1)
+    test.reset_index(inplace=True)
+    return train.iloc[:, 1:], validate.iloc[:, 1:], test.iloc[:, 1:]
+\end{lstlisting} 
+\pagebreak
+
+\section{Python Class for Network Layers}
+The following code shows the creation of the \verb|NeuronLayer| class. The full subroutine can be found in the included file \verb|NLObjects.py|.
+
+\begin{lstlisting}[style = Python]
+class NeuronLayer:
+    # define attributes
+    def __init__(self, r, s, f, j):
+        self.r = r # Number of inputs
+        self.s = s # number of neurons
+        self.f = f # Transfer function
+        self.j = j # derivatve transfer function
+
+    def setWeightBias(self, w, b):
+        self.w = w  # set weights
+        self.b = b # set biases
+        self.r = w.shape[1]
+        self.s = w.shape[0]
+        return
+
+    def FP(self, p):
+        self.p = p
+        self.n = self.w * self.p + self.b
+        self.a = self.f(self.n)
+        return
+
+    def update(self, learning_rate, sensitivity):
+        self.w = self.w - learning_rate * sensitivity * self.p.T
+        self.b = self.b - learning_rate * sensitivity
+        return
+\end{lstlisting} 
+\pagebreak
+
+\section{Initializing the Network Architecture in Python}
+This code shows the process used to instantiate layers and initialize the parameters of the network.
+
+\begin{lstlisting}[style = Python]
+# --------------------- specify network architecture ---------------------
+num_neurons1 = 10  # layer 1
+num_neurons2 = 2   # layer 2
+alpha = 0.01       # learning rate
+epoch = 100        # number of iterations
+nlayer1 = nl.NeuronLayer(num_neurons1, num_inputs, nl.tansig, nl.j_tansig)      # instantiate layer 1
+nlayer2 = nl.NeuronLayer(num_neurons2, num_neurons1, nl.softmax, nl.j_softmax)  # instantiate layer 2
+np.random.seed(0)
+# randomly initialize weight and bias on the interval [-0.5, 0.5]
+W1 = np.matrix(np.random.rand(num_neurons1, num_inputs) - 0.5)
+b1 = np.matrix(np.random.rand(num_neurons1, 1) - 0.5)
+W2 = np.matrix(np.random.rand(num_neurons2, num_neurons1) - 0.5)
+b2 = np.matrix(np.random.rand(num_neurons2, 1) - 0.5)
+# initialize cross entropy loss (training & validation)
+ce_t = []
+ce_v = []
+global s1_all, s2_all
+# pass on randomly initialized weights and biases to the network
+nlayer1.setWeightBias(w=W1, b=b1)
+nlayer2.setWeightBias(w=W2, b=b2)
+\end{lstlisting} 
+\pagebreak
+
+\section{Early Stopping}
+The code below shows the early stopping condition used to prevent the model from overfitting.
+
+\begin{lstlisting}[style = Python]
+if j == 0:
+	val_fail = []
+elif ce_v[j] > ce_v[j-1]:
+	val_fail.append(1)
+	if len(val_fail) == 5:
+		print 'Validation error has increased for 5 consecutive epochs. Early stopping at epoch {}'.format(j)
+		break
+else:
+	val_fail = []
+\end{lstlisting}
 

### 2017-06-30 Python Code/Split.py

@@ -3,25 +3,20 @@ import numpy as np
 
 def split(dataset, num_inputs, t, v):
     # normalize dataset into range between -1 and 1
-
     norm = dataset.ix[:, slice(0,num_inputs)].apply(lambda x: -1 + 2 * (x - x.min()) / (x.max() - x.min()), axis=0)
     dataset = pd.concat([norm, dataset.ix[:,'window':]], axis=1)
-
     # split dataset by target class
     dataset0 = dataset.ix[dataset.ix[:, 'window'] == 0, :]
     dataset1 = dataset.ix[dataset.ix[:, 'window'] == 1, :]
-
     # build datasets by target class
     train0, validate0, test0 = np.split(dataset0.sample(frac=1),
                                         [int(t * len(dataset0)), int((1 - v) * len(dataset0))])
     train1, validate1, test1 = np.split(dataset1.sample(frac=1),
                                         [int(t * len(dataset1)), int((1 - v) * len(dataset1))])
-
     # re-combine by target class
     train = pd.concat([train0, train1])
     validate = pd.concat([validate0, validate1])
     test = pd.concat([test0, test1])
-
     # shuffle order and reset index numbers
     train = train.sample(frac=1)
     train.reset_index(inplace=True)
@@ -29,5 +24,4 @@ def split(dataset, num_inputs, t, v):
     validate.reset_index(inplace=True)
     test = test.sample(frac=1)
     test.reset_index(inplace=True)
-
     return train.iloc[:, 1:], validate.iloc[:, 1:], test.iloc[:, 1:]

### 2017-06-30 Python Code/main_code.py

@@ -1,140 +1,0 @@
-# ----------------------------------------------------------------------------------------------------------------------
-# Import statements
-# ----------------------------------------------------------------------------------------------------------------------
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-import NLObjects as MLP
-from DataPrep import GlassImport
-from Split import split
-# ----------------------------------------------------------------------------------------------------------------------
-# Data pre-processing
-# ----------------------------------------------------------------------------------------------------------------------
-# import the glass dataset
-glass = GlassImport()
-# return the dimension of each input, R
-dim_inputs = glass.shape[1] - 2
-# split into training, validation, & testing sets
-train, validate, test = split(glass, dim_inputs, t=0.7, v=0.15)
-# number of neurons in layer 1
-num_neurons1 = 10
-# number of neurons in layer 2
-num_neurons2 = 2
-# learning rate
-alpha = 0.1
-# num_iter
-epoch = 150
-# instantiate layer 1
-nlayer1 = MLP.NeuronLayer(num_neurons1, dim_inputs, np.tanh, MLP.j_tansig)
-# instantiate layer 2
-nlayer2 = MLP.NeuronLayer(num_neurons2, num_neurons1, MLP.softmax, MLP.j_softmax)
-# set seed for reproducible results
-np.random.seed(150)
-# randomly initialize weight and bias on the interval [-0.5, 0.5]
-W1 = np.matrix(np.random.rand(num_neurons1, dim_inputs) - 0.5)
-b1 = np.matrix(np.random.rand(num_neurons1, 1) - 0.5)
-W2 = np.matrix(np.random.rand(num_neurons2, num_neurons1) - 0.5)
-b2 = np.matrix(np.random.rand(num_neurons2, 1) - 0.5)
-# pass the randomly initialized weight and bias onto the network object
-nlayer1.setWeightBias(w=W1, b=b1)
-nlayer2.setWeightBias(w=W2, b=b2)
-# ------------------- training set -------------------------
-# matrix of inputs
-p_train = train.iloc[:, range(dim_inputs)].T
-p_train = np.asmatrix(p_train)
-# matrix of targets
-target = train.iloc[:, -2:]
-target = np.asmatrix(target)
-# ------------------- validation set -------------------------
-# matrix of inputs
-p_val = validate.iloc[:, range(dim_inputs)].T
-p_val = np.asmatrix(p_val)
-# matrix of targets
-target_val = validate.iloc[:, -2:]
[…]

### 2017-06-30 Paper/paper.tex
Experimental Setup

@@
 
+\subsection{Testing the Network}
+Testing the network is a simple process since the network is trained and we just need to feed the inputs in to the network.
+Since the transfer function in the output layer is the \emph{softmax} function, the network output is the class probabilities between 0 and 1.
+Once the class probabilities are predicted, we need to use these to convert the probabilities to class predictions.
+To do this, we use 0.5 as the cutoff, so if one neuron receives a higher probability than the other, that one is chosen to be the positive output from the network and the other is assigned to be null.
+Then, we compare the predictions to the target which tells us the true class.
+Using this information, we can calculate classification error, accuracy, true and false positive rates, and construct a confusion matrix.
+We use all of these as performance metrics in discuss them in the next section.
+The code used to test the network can be found in Appendix F.
+
 \section{Results}
@@
 \pagebreak
+
 \appendix
@@
 \end{lstlisting} 
+\pagebreak
+
+\section{Testing the Network}
+The code below shows the code used to test the network and calculate the performance metrics.
+
+\begin{lstlisting}[style = Python]
+# --------------------- confusion matrix ---------------------
+actual = pd.Series(test.iloc[:, -2], name='Actual')  # actual values (targets)
+input = np.matrix(test.iloc[:, slice(0, num_inputs)]).transpose()  # network inputs, p
+nlayer1.FP(input)  # layer 1 net-input
+nlayer2.FP(nlayer1.a)  # layer 2 net-input
+predict = nl.classify(nlayer2.a)  # predicted values from network
+predict = np.array(predict).flatten()
+predict = pd.Series(predict, name='Predicted')
+confusion = pd.crosstab(actual, predict, margins=False)  # create confusion matrix
+confusion = confusion.astype(float)  # convert values to floats
+print confusion  # output confusion matrix to the console
+# --------------------- accuracy metrics ---------------------
+ERR = (confusion.loc[0, 1] + confusion.loc[1, 0]) / len(predict)
+ACC = 1 - ERR
+FPR = confusion.iloc[0, 1] / (confusion.iloc[0, 1] + confusion.iloc[0, 0])
+TPR = confusion.iloc[1, 1] / (confusion.iloc[1, 0] + confusion.iloc[1, 1])
+\end{lstlisting}
 

### 2017-06-30 Python Code/NLObjects.py

@@ -1,5 +1,7 @@
 import numpy as np
-# create neuron class
+# ----------------------------------------------------------------------------------------------------------------------
+# Create Class for Network Layers
+# ----------------------------------------------------------------------------------------------------------------------
 class NeuronLayer:
     # define attributes
     def __init__(self, r, s, f, j):
@@ -25,7 +27,9 @@ class NeuronLayer:
         self.w = self.w - learning_rate * sensitivity * self.p.T
         self.b = self.b - learning_rate * sensitivity
         return
-
+# ----------------------------------------------------------------------------------------------------------------------
+# transfer functions
+# ----------------------------------------------------------------------------------------------------------------------
 # hardlimit transfer function a = hardlim(n)
 def hardlim(n):
     if n >= 0:
@@ -33,39 +37,29 @@ def hardlim(n):
     else:
         a = 0
     return a
-
-
 # log-sigmoid transfer function a = logsig(n)
 def logsig(n):
     return 1 / (1 + np.exp(-n))
-
-
 # linear transfer function
 def purelin(n):
     return n
-
 # hyperbolic tangent sigmoid transfer function
 def tansig(n):
     e_pos = np.exp(n)
     e_neg = np.exp(-n)
     return (e_pos - e_neg) / (e_pos + e_neg)
-
 # softmax transfer function a = e^n/sum(e^n)
 def softmax(n):
     return np.exp(n) / np.sum(np.exp(n), axis=0)
-
-# classify output
-def classify(a):
-    c = np.where(a[0] >= 0.5, 0, 1)
-    return c
-
+# ----------------------------------------------------------------------------------------------------------------------
+# transfer function gradients
+# ----------------------------------------------------------------------------------------------------------------------
 # tansig derivative jacobian matrix
 def j_tansig(a):
     tg_diags = 1 - np.power(np.tanh(a),2)
     tg = np.zeros((a.shape[0], a.shape[0]))
     np.fill_diagonal(tg, tg_diags)
     return tg
-
 # softmax derivative jacobian matrix
 def j_softmax(a):
     jacobian = np.zeros((a.shape[0], a.shape[0]))
@@ -76,30 +70,38 @@ def j_softmax(a):
             else:
                 jacobian[i, j] = -a[i] * a[j]
     return jacobian
-
-
+# ----------------------------------------------------------------------------------------------------------------------
+# Performance Index
+# ----------------------------------------------------------------------------------------------------------------------
[…]

### 2017-06-30 Python Code/neuron.py

@@ -1,263 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-# create neuron class
-class Neuron:
-
-    # define attributes
-    def __init__(self, input, weight, bias):
-        self.input = input
-        self.weight = weight
-        self.bias = bias
-
-    # methods
-
-    # calculate net input n = Wp + b
-    def net_input(self):
-        n = self.weight * self.input + self.bias
-        return n
-
-    # hardlimit transfer function a = hardlim(n)
-    def hardlim(self):
-        if self.net_input() >= 0:
-            a = 1
-        else:
-            a = 0
-        return a
-
-    # log-sigmoid transfer function a = logsig(n)
-    def logsig(self):
-        a = 1/(1 + np.exp(-self.net_input()))
-        return a
-
-    # linear transfer function
-    def purelin(self):
-        a = self.net_input()
-        return a
-
-    # softmax transfer function a = e^n/sum(e^n)
-    def softmax(self):
-        #a = np.exp(self.net_input())/np.exp(self.net_input()).sum()
-        a = np.exp(self.net_input()) / np.exp(self.net_input()).sum(axis=0)
-        return a
-
-    # softmax derivative jacobian matrix
-    def j_softmax(self,a,num):
-        jacobian = np.empty([num, num])
-        for i in range(num):
-            for j in range(num):
-                if i == j:
-                    jacobian[i, j] = a[i] * np.sum(a - a[i])
-                else:
-                    jacobian[i, j] = -a[i] * a[j]
-        return jacobian
-
-    # hyperbolic tangent sigmoid transfer function
-    def tangsig(self):
-        a = 2/(1 + np.exp(-2 * self.net_input())) - 1
-        return a
-
-    # tansig derivative jacobian matrix
-    def j_tansig(self,a,num):
-        jacobian = np.diag(np.ones(num)) - np.diagflat(np.power(a, 2))
-        return jacobian
-
-    # classify output
-    def classify(self,a):
-       c = np.where(a[0]>=0.5,1,0)
-       return c
-
-# define cross entropy error calculation function for softmax (assuming targets of 1,0)
-def cross_entropy(a,t):
-    i = np.where(a==0)
-    a[i] = 1e-15 # replace 0s so ln doesn't produce infinity
-    e = -t * np.log(a)
-    return e
-
[…]

### 2017-07-02 Paper/paper.tex
Paper edits

@@
 
+Figure 2 below shows the $\log$ cross-entropy loss of the training and validation sets while we are training the network.
+We can see no clear signs of overfitting as the $\log$ of the loss function continues to decrease monotonically for both sets over time.
+If the $\log$ of the loss function for the validation set started to increase over time, then this would be an indication of overfitting, but our early stopping condition would stop the network from training any further.
+Since the $\log$ of the loss function has not started to level off after 100 epochs at a learning rate of 0.01, we may be able to increase the number of epochs or increase the learning rate to see if this yields better results.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width = 5in]{figs/CEloss.png}
+\caption{$\log$ Cross-Entropy Loss on Training and Validation Sets over Time}
+\end{figure}
+
 \subsection{Testing the Network}
@@
 
+For pattern recognition problems, the confusion matrix is one of the most widely used tools to visualize the misclassification of a network.
+The columns of a confusion matrix represent the target class and the rows represent the predicted class.
+Thus, the correctly classified inputs are on the diagonal cells of the confusion matrix, whereas the off-diagonal cells show misclassified inputs.
+The confusion matrix from the test set can be seen in Figure 3 below.
+\begin{figure}[H]
+\centering
+\includegraphics[width = 3in]{figs/confmat.png}
+\caption{Confusion Matrix for the Multi-Layer Perceptron Network on the Test Set}
+\end{figure}
+
+Recall that there are 33 samples in our test set.
+Since the the diagonal cells add up to 32, our model correctly classified 32 out of the 33 samples.
+We can use this information to calculate the accuracy of our model:
+\begin{equation}
+Accuracy = \frac{TP + TN}{FP + FN + TP + TN}.
+\end{equation}
+The top left cell of the confusion matrix represents true negatives (TN) and the bottom right cell represents true positives (TP).
+Adding up the diagonal cells and dividing by the total number of samples, we find that our model has a 97\% accuracy.
+This is higher than the vast majority of the tests we ran in MATLAB.
+We can use this accuracy score to find that our misclassification error is 3\% since:
+\begin{equation}
+accuracy = 1 - error \implies error = 1 - accuracy. 
+\end{equation}
+However, the error can also be computed as
+\begin{equation}
+error = \frac{FP + FN}{FP + FN + TP + TN}
+\end{equation}
+which is essentially the sum of the off-diagonal cells divided by the sample size.
+
+In addition to simply looking at the accuracy and error of the model on the training set, we can break this down even further since our classes are imbalanced.
+From the confusion matrix in Figure 3, we can see the positive class contains more than \emph{three} times as many elements as the null class.
+The false positive and true positive rates can be computed as follows:
+\begin{equation}
+FPR = \frac{FP}{FP + TN} , 
+\end{equation}
+\begin{equation}
+TPR = \frac{TP}{FN + TP}.
+\end{equation}
+We find that the false positive rate and true positive rate is 0 and 0.96, respectively.
+A false positive rate of zero implies there was not a single time that our model committed a \emph{Type I} error.
+That is, the network did not predict a glass sample to be a a window glass when it was a non-window glass.
+This is very important for use in forensic analysis as it is costly to tie someone to a crime scene when they were not actually there.
+In fact, this could potentially result in a wrongful conviction for a suspect if they are unlucky.
+However, it is equally as important not to commit \emph{Type II} errors in this scenario.
+
+Type II errors correspond to false negatives.
+That is, the model predicts a sample of glass is a non-window glass when it is window glass.
+From the bottom left cell in Figure 3, we can see our model produced one false negative.
+Type II errors are costly in the sense that if we classify a window glass as a non-window glass in forensic analysis, we could wrongly-free a suspect back into society when they in fact committed the crime.
+However, we should not be alarmed at the single false negative as the true positive rate is 0.96.
+This means that 96\% of the time, we are correctly classifying the positive class since the true positive rate is the number of true positives divided by the total number of positive targets.
+
+Precision and recall are two other performance metrics related to true positive and false positive rates. Precision can be denoted by 
+\begin{equation}
+PRE = \frac{TP}{TP + FP}
+\end{equation}
+and recall is analogous to the true positive rate.
+Our precision happens to be 100\% which is very good.
+It is important for a model to be both accurate and precise as accuracy is the model being correct and precision is the actual reproducibility of our measurements.
+If a model is accurate and not precise, it can fluctuate drastically between tests with regard to the performance metrics.
+Table 1 below gives a summary of the scoring metrics.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+\multicolumn{2}{|l|}{Summary of Scoring Metrics} \\ \hline
+Accuracy                       & 97\%            \\ \hline
+Error                          & 3\%             \\ \hline
+False Positives                & 0               \\ \hline
+False Positive Rate            & 0\%             \\ \hline
+True Positives                 & 24              \\ \hline
+True Positive Rate             & 96\%            \\ \hline
+Precision                      & 100\%           \\ \hline
+Recall                         & 96\%            \\ \hline
+\end{tabular}
+\caption{Summary of Scoring Metrics used on Test Set}
+\label{my-label}
+\end{table}
+
 \section{Conclusion}
 \label{sec7}
+
+Using a multi-layer perceptron, we were able to train a neural network that correctly classifies whether a glass sample came from a window or not 97\% of the time.
+We used several optical and physicochemical properties of glass to make these predictions.
+We quickly implemented the multi-layer perceptron in MATLAB's Neural Network Toolboox (\verb|nntool|) and then replicated the network in Python.
+
+This project helped us to understand how flexible and powerful two-layer neural networks can be.
+Almost any problem can be solved with a two-layer neural network if you have good data, preprocess the data in accordance with the selected network architecture, and choose the right transfer functions and performance index to solve the problem.
+In class, all of the back propagation problems we solved by hand only had one neuron in the output layer and used a \emph{purelin} transfer function.
+Building this network in Python without using packages taught us how to use backpropagation with a different performance index and how to derive the sensitivity for the output layer with respect to the output layer and the hidden layer.
+
+We feel that there are several ways the model could be improved, aside from the general performance.
+The code we wrote in Python is very rough and can be neatly packaged together if more time is spent working on the class we defined.
+Moreover, we could add the backpropagation process as a method to the class instead of doing this outside of the class structure as we do in the \verb|NeuronLayer.py| file.
+In the future, we wish to add the ability to pass a performance index as input to the class when instantiating.
+
+Similarly, we feel there are a few more performance metrics we could use to interpret the results of the model.
+A receiver operator characteristic (ROC) curve would allow us to see how changing the decision threshold would impact the false positive and true positive rates.
+This is especially important if we need to bias the model towards one class in the case of imbalanced classes.
+After plotting the ROC curve, we could compute the area under the curve (AUC) to characterize the performance of the model.
+
+The one thing our model is lacking is a good way to compare it to the results from the MATLAB model.
+In the future, we may wish to find a way to split our training, validation, and test sets up ahead of time and then pass each programming framework the same sets.
+This would give an apples-to-apples comparison of the models as they would be trained, validated, and tested on the exact same data.
 

### 2017-07-02 Paper/paper.tex
Incorporating Mark's Edits

@@
 It is difficult to know before we train a neural network if we have enough data, but the amount of data required is directly related to the complexity of the underlying decision boundary we are trying to implement.
-We won't
-know how complex the decision boundary we are trying to approximate is until we train the network, but we feel confident using the dataset as many others have used the dataset and found robust results.
+We won't know how complex the decision boundary we are trying to approximate is until we train the network, but we feel confident using the dataset as many others have used the dataset and found robust results.
 Several other papers in the literature use much more complex methods than we will employ and have not found the size of the data to be an issue.
 
-We have chosen a two-layer perceptron network with tangent-sigmoid transfer functions in the hidden layer and \emph{softmax} transfer functions in the output layer.
+We have chosen a two-layer perceptron network with \emph{tansig} transfer functions in the hidden layer and \emph{softmax} transfer functions in the output layer.
 This is a fairly standard network for pattern recognition.
@@
 First, we will use the Neural Network Toolbox, specifically the Neural Network Pattern Recognition Tool (\verb|nprtool|) train, validate, and test our network.
-We use this framework to start with a simple graphical user interface to quickly ensure our specified network architecture is appropriate and to get baseline performance statistics.
+This framework allows us to quickly ensure our specified network architecture is appropriate and to get baseline performance statistics.
 Then, we will replicate the analysis in Python to gain practical experience building network architectures in a scripting language.
@@
 \label{sec4}
-Once the data is preprocessed, the next step is to decide on or create a network architecture.
+Once the data is preprocessed, the next step is to create a network architecture.
 The basic network architecture is determined by the problem we wish to solve.
@@
 If the results of the network are unsatisfactory after training and testing with one hidden layer, we will retrain with an additional hidden layer, but we do not anticipate having to do this.
-The \emph{tansig} transfer function is usually preferred to the \emph{tansig} transfer function in the hidden layers since it produces outputs (which are inputs to the next layer) that are centered near zero, whereas the \emph{tansig} transfer function always produces positive outputs.
+The \emph{tansig} transfer function is usually preferred to the \emph{logsig} transfer function in the hidden layers since it produces outputs (which are inputs to the next layer) that are centered near zero, whereas the \emph{logsig} transfer function always produces positive outputs.
 
@@
 \label{sec4.2.3}
-Once we have propagated the input forward, we tcompare the network output to the targets so we can use the error to adjust the weights and biases accordingly.
+Once we have propagated the input forward, we compare the network output to the targets so we can use the error to adjust the weights and biases accordingly.
 But, in the case of a multi-layer network, the errors and the performance index we use to evaluate those errors are no longer just a function of the weights.
@@
 \frac{\partial n^{m+1}_{i}}{\partial \mathbf{n}^{m}_{j}} &= \frac{\partial \bigg(\sum_{l=1}^{S^{m}} w_{i, l}^{m+1}a_{l}^{m} + b_{i}^{m+1}\bigg)}{\partial n_{j}^{m}} = w_{i, j}^{m+1} \frac{\partial a_{j}^{m}}{\partial n_{j}^{m}} \\
-&= w_{i, j}^{m+1} \frac{\partial f^{m}(n_{j}^{m}}{\partial n_{j}^{m}} = w_{i, j}^{m+1} \dot{f}^{m}(n_{j}^{m}) ,  
+&= w_{i, j}^{m+1} \frac{\partial f^{m}(n_{j}^{m})}{\partial n_{j}^{m}} = w_{i, j}^{m+1} \dot{f}^{m}(n_{j}^{m}) ,  
 \end{split}
@@
 Once the class probabilities are predicted, we need to use these to convert the probabilities to class predictions.
-To do this, we use 0.5 as the cutoff, so if one neuron receives a higher probability than the other, that one is chosen to be the positive output from the network and the other is assigned to be null.
+To do this, we use 0.5 as the cutoff, so if one neuron receives a higher probability than the other, that one is chosen to be the positive output from the network and the other is assigned to be zero.
 Then, we compare the predictions to the target which tells us the true class.
@@
 Almost any problem can be solved with a two-layer neural network if you have good data, preprocess the data in accordance with the selected network architecture, and choose the right transfer functions and performance index to solve the problem.
-In class, all of the back propagation problems we solved by hand only had one neuron in the output layer and used a \emph{purelin} transfer function.
+In class, all of the backpropagation problems we solved by hand only had one neuron in the output layer and used a \emph{purelin} transfer function.
 Building this network in Python without using packages taught us how to use backpropagation with a different performance index and how to derive the sensitivity for the output layer with respect to the output layer and the hidden layer.
